### 2017-03-09 PaperEx170308.Rmd
paper examples

---
title: "Mixture Modeling Paper Results"
author: "Sarah Sullivan"
date: "March 8, 2017"
output: word_document
---

```{r loadfunctions}
library(mixsmsn)
library(sn)
#-----------------------------------------------
#[IC] Load Functions
#-----------------------------------------------

#-----------------------------------------------
#[IC1]  fitloops
#-----------------------------------------------

#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skewnormal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
#accepts a dataframe with columns id and data
fitloops<-function(datawithids,loops=10,maxcp=5){
  dists="Both"
  datawithids<-unique(datawithids)
  data<-datawithids$data
  datapts<-length(data)
  norms<-vector("list")
  errors<-matrix(0, nrow=maxcp, ncol=2)
  loopErr = function(data,i,j,family) {
    tryCatch(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5)),
             error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
               NA}) 
  }

  for (i in 1:maxcp){
    norms$mc[[i]]<-vector("list", maxcp)
    norms$ms[[i]]<-vector("list", maxcp)
    j=1
    while (max(errors)<=2 & j<=loops) {

      norms$mc[[i]][[j]]<-vector("list", 15)
      norms$ms[[i]][[j]]<-vector("list", 15)
      if (dists=="Both") {
        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")  
        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
      } else if (dists=="Normal"){ 
        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")
      } else if (dists=="Skew.normal"){
        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
      } else {
        stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew.normal'"))
      }

      message(paste(i, "Component(s), Loop", j))
      if (is.na(norms$mc[[i]][[j]][1])==T) {
        norms$mc[[i]][[j]]<-vector("list")
        norms$mc[[i]][[j]]$bic<-NA
        errors[i,1]<-errors[i,1]+1
      } 
      if (is.na(norms$ms[[i]][[j]][1])==T) {
        norms$ms[[i]][[j]]<-vector("list")
        norms$ms[[i]][[j]]$bic<-NA
        errors[i,2]<-errors[i,2]+1
      } 
      if (dists=="Normal"){ 
        norms$ms[[i]][[j]]<-vector("list")
        norms$ms[[i]][[j]]$bic<-NA
      } else if (dists=="Skew.normal"){
        norms$mc[[i]][[j]]<-vector("list")
        norms$mc[[i]][[j]]$bic<-NA
      }
      j=j+1
      ic=i
    }
  }

  if (i==maxcp & j==loops+1 & max(errors[maxcp, 2])<=2 & max(errors[maxcp, 1])<=2 ) {
    nobj<-vector("list")
    nobj$mc<-matrix(nrow=maxcp, ncol=loops)
    nobj$ms<-matrix(nrow=maxcp, ncol=loops)

    for (i in 1:maxcp){
      for (j in 1:loops){
        nobj$mc[i,j]<-norms$mc[[i]][[j]]$bic
        nobj$ms[i,j]<-norms$ms[[i]][[j]]$bic
      }
    }

    for (i in 1:maxcp){
      if (dists=="Both" | dists=="Normal"){
        nobj$lmin$mc[i]<-which.min(nobj$mc[i,])
      } else {
        nobj$lmin$mc[i]<-1
      }
      if (dists=="Both" | dists=="Skew.normal"){
        nobj$lmin$ms[i]<-which.min(nobj$ms[i,])
      } else {
        nobj$lmin$ms[i]<-1
      }
    }

    fitres<-vector("list")
    for (i in 1:maxcp){
      fitres$no[[i]]<-vector("list", 15)
      fitres$sn[[i]]<-vector("list", 15)
      fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
      fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
    }
    fitres$loops<-loops
    fitres$maxcp<-maxcp
    fitres$datawithids<-datawithids
    fitres$errors<-errors
    return(fitres)
  } else {
    if (which.max(errors)<=maxcp){
      dist<-"Normal"
    } else {
      dist<-"Skew.normal"
    }
    if (ic<=2){
      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s). First ensure that the necessary packages have been loaded correctly. If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skewnormal or normal distribution assumptions."))
    } else {
      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s). Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
    }
  }  
}

#-----------------------------------------------
#[IC2]  bestfits
#-----------------------------------------------
#Determines the optimal number of distributions and type of distributions (as determined by BIC), was well as the other indicated functions. Also creates table of BICs which can be fed into the bicgraph function to visualize which combination of distributions and component numbers is optimal by BIC.

bestfits<-function(fitres){

  maxcp<-fitres$maxcp
  loops<-fitres$loops
  nobj<-vector("list")

  for (i in 1:maxcp){
    for (j in 1:loops){
      nobj$no[i]<-fitres$no[[i]]$bic
      nobj$sn[i]<-fitres$sn[[i]]$bic
    }
  }

  nobj$minind<-data.frame(no=as.numeric(c(NA)), sn=as.numeric(c(NA)))

  if (is.na(nobj$no[1])==T){
    nobj$minind$no<-1
  } else {
    nobj$minind$no <- (which.min(nobj$no))
  }

  if (is.na(nobj$sn[1])==T){
    nobj$minind$sn<-1
  } else {
    nobj$minind$sn <- (which.min(nobj$sn))
  }

  nobj$min<-data.frame(type=c("no", "sn"), bic=as.numeric(c(NA, NA)))

  nobj$min$type<-as.character(c("no", "sn"))
  type<-c("no", "sn")
  nobj$min$type<-type

  nobj$min[1,2]<-fitres$no[[nobj$minind$no[1]]]$bic
  nobj$min[2,2]<-fitres$sn[[nobj$minind$sn[1]]]$bic

  nobj$distnum<-which.min(nobj$min$bic)
  nobj$dist<-nobj$min$type[nobj$distnum]

  nobj$best<-fitres[[nobj$dist]][[nobj$minind[nobj$dist][1,1]]]

  if (nobj$dist=="no"){
    nobj$distp="Normal"
    nobj$distf="Normal"
  } else if (nobj$dist=="sn") {
    nobj$distp="Skew normal"
    nobj$distf="Skew.normal"
  } else {
    nobj$distp=""
  }

  model<-vector("list")
  model$sn2<-fitres$sn[[2]]
  model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
  model$norm2<-fitres$no[[2]]
  model$normbest<-fitres$no[[nobj$minind$no[1]]]
  model$best<-nobj$best
  model$bictab<-data.frame(normal=c(NA)[rep(c(1), times=maxcp)], skew.normal=c(NA)[rep(c(1), times=maxcp)])
  model$bictab$normal<-nobj$no
  model$bictab$skew.normal<-nobj$sn
  componentnames<-vector()
  for (i in 1:maxcp){
    componentnames[i]<-paste(i, " Component(s):  ", sep="")  
  }    
  rownames(model$bictab)<-componentnames
  colnames(model$bictab)<-c("Normal", "Skew normal")
  model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
  rownames(model$summary)<-c("Best Overall", "Best Skew normal  ", "Best Normal", "Two Skew normal", "Two Normal")
  colnames(model$summary)<-c("Description         ", "BIC  ")

  model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
  model$datawithids<-fitres$datawithids

  print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
  print("")
  print("Bayesian Information Criterion (BIC) Matrix")
  print("BIC should be minimized and a difference of 10 BIC indicates strong evidence that the model with lower BIC is superior")
  print(model$bictab)
  print("Below is a table of the BIC's of the most common distribution and number of component combinations to base a cutpoint on")
  print(model$summary)

  return(model)

}

#-----------------------------------------------
#[IC3]  bicgraph
#-----------------------------------------------
#Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.

bicgraph<-function(model,title="BIC by type and number of distributions",setcolor=c("magenta", "blue", "black")){
  color<-setcolor
  if (is.na(model$bictab$Normal[1])==F) {
    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n")
    lines(model$bictab$`Skew normal`, type="o", pch=2, lty=2, col=color[2])
  } else {
    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n")
    lines(model$bictab$`Skew normal`, pch=1, col=color[1], type="o")    
  }  
  points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
  axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))

}

#-----------------------------------------------
#[IC4]  fitpick
#-----------------------------------------------
#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.

fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
  if (dist=="" & is.na(ncomp)==T) {
    dist=bestfits$bestdesc$dist
    ncomp=bestfits$bestdesc$ncomp
  }
  if (ncomp==1) {
    stop("Single component specified so there is no cut-point")
  }

  data=fitres$datawithids$data
  n=ncomp
  if (dist!="Normal" & dist!="Skew.normal") {
    stop("Distribution must be specified as either 'Normal' or 'Skew.normal'")
  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew.normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
    stop("Must choose a distribution and number of components that were fitted in the fitloops function")
  } else {
    if (dist=="Normal"){
      dist="no"
    } else {
      dist="sn"
    }

    if (dist=="no"){
      distp="Normal"
    } else if (dist=="sn") {
      distp="Skew Normal"
    } else {
      distp=""
    }

    singlefit<-vector("list")
    singlefit<-fitres[[dist]][[ncomp]]

    comp<-data.frame(ogroup=singlefit$group, data=data)
    comp$index <- as.numeric(row.names(comp))
    comp <- comp[order(comp$data),] 
    comp$diff<-c(0, diff(comp$ogroup))
    nl=as.numeric(dim(comp)[1])

    #Will recode the distributions so that they are in order with 1 being the lowest
    comp$c<-vector(mode="numeric", length=nl)
    comp$c[1]<-1
    for (i in 2:nl){
      if (comp$diff[i]==0) {
        comp$c[i]=comp$c[i-1]
      } else {
        comp$c[i]=comp$c[i-1]+1
      }
    }
    comp <- comp[order(comp$index),]
    singlefit$group2<-comp$c
    comp <- comp[order(comp$data),]

    intDistErr = function() {
      tryCatch(data.frame(c=unique(comp$c), o=unique(comp$ogroup)),
               error = function(e) {data.frame(c=c(NA)[rep(c(1), times=length(comp$c))], o=c(NA)[rep(c(1), times=length(comp$c))])}) 
    }

    cross<-intDistErr()
    if (suppressWarnings(is.na(cross[1,1])==T)){
      stop("This combination of distribution and number of components yeilds interrupting distributions, which make determining a cut-point impossible. Please choose a different combination of distribution and number of components")
    }

    for (i in 1:n){
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
      singlefit$sigma2[i]<-fitres[[dist]][[ncomp]]$sigma2[cross$o[i]]
      singlefit$shape[i]<-fitres[[dist]][[ncomp]]$shape[cross$o[i]]
      singlefit$pii[i]<-fitres[[dist]][[ncomp]]$pii[cross$o[i]]
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
    }

    singlefit$obs.prob<-NULL

    #Detect if there are any interupting distributions (can also be recoded to kick someone out and tell them to pick something else:
    if (max(comp$c)>max(comp$ogroup)) {
      print("Warning: there is at least one interupting distribution, which may make the determination of a cutoff and the calculation of an indeterminate range difficult in later functions. If later functions fail, try a different combination of distribution and number of components (ideally the next most optimal by BIC).")
    }

    par<-vector("list")
    par<-data.frame(mean=c(NA)[rep(c(1), times=n)], sd=c(NA)[rep(c(1), times=n)], shape=c(NA)[rep(c(1), times=n)], pii=c(NA)[rep(c(1), times=n)])
    for (i in 1:n){
      par$mean[i]=singlefit$mu[i]
      par$sd[i]=sqrt(singlefit$sigma2[i])
      par$sigma2[i]=singlefit$sigma2[i]
      par$shape[i]=singlefit$shape[i]
      par$pii[i]=singlefit$pii[i]
    }

    parameters<-vector("list")
    parameters$singlefit<-singlefit
    parameters$par<-par
    parameters$datawithids<-fitres$datawithids
    parameters$desc<-vector("list")
    parameters$desc$ncomp<-ncomp
    parameters$desc$dist<-distp
  }

  return(parameters)
}

#-----------------------------------------------
#[IC5]  uncert
#-----------------------------------------------
#Performs uncertainty calculations.

uncert<-function(fitpickobj){

  data<-fitpickobj$datawithids$data
  par<-fitpickobj$par
  fit<-fitpickobj$singlefit
  n=fitpickobj$desc$ncomp
  #Create Uncertainties
  v<-c(seq(min(data),max(data),0.0001))

  nl<-length(v)
  PD<-vector(mode="numeric", length=n)
  PD[1]<-1
  g1<-0
  g2<-0
  Pv.D<-matrix(0, nl, n)
  Pv.DPD<-matrix(0, nl, n)
  PDT<-matrix(0, nl, 1)
  PD.v<-matrix(0, nl, n)
  for (i in 2:n) {
    PD[i]<-par$pii[i]
    PD[1]<-PD[1]-PD[i]
  }
  for (i in 1:n){
    if (fitpickobj$desc$dist=="Normal"){
      Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
    } else {
      Pv.D[,i]<-dsn(v,xi=par$mean[i],omega=par$sd[i],alpha=par$shape[i])
    }
  }
  for (i in 1:n){
    for (j in 1:nl){
      Pv.DPD[j,i]<-Pv.D[j,i]*PD[i]
    }
  }
  PDT<-rowSums (Pv.DPD)
  for (i in 1:n){
    for (j in 1:nl){
      PD.v[j,i] <- Pv.DPD[j,i] / (PDT[j])
    }
  }

  mcluncert<-1-apply(PD.v, 1, max) 
  uncertmat<-PD.v
  restrmat<-cbind(PD.v, v)
  uncertout<-vector("list")
  uncertout$uncertmat<-uncertmat
  uncertout$mcluncert<-mcluncert
  uncertout$restrmat<-restrmat
  uncertout$singlefit<-fit
  uncertout$v<-v
  uncertout$datawithids<-fitpickobj$datawithids
  uncertout$desc<-fitpickobj$desc
  return(uncertout) 
}

#-----------------------------------------------
#[IC6]  multcut
#-----------------------------------------------
#Calculates raw percentages and counts positive and negative for all possible cut points between distributions
multcut<-function(uncertobj){

  PD.v<-uncertobj$uncertmat
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp
  ncomp=n
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  g1<-vector("list")
  g2<-vector("list")
  uncertdf<-vector("list")
  uncertainty.v<-vector("list")
  cutpoint<-c(0)[rep(c(1), times=ncomp-1)]
  for (j in 1:(ncomp-1)){
    g1[[j]]<-0
    g2[[j]]<-0
    for (i in 1:j){
      g1[[j]]<-g1[[j]]+PD.v[,i]
    }
    for (i in (j+1):n){
      g2[[j]]<-g2[[j]]+PD.v[,i]
    }
    uncertainty.v[[j]]<-1-pmax(g1[[j]], g2[[j]])
  }

  for (j in 1:(ncomp-1)){
    uncertdf[[j]]<-data.frame(v=uncertobj$v, uncertainty.v=uncertainty.v[[j]])
    cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
  }
  class2<-vector("list")
  for (j in 1:(ncomp-1)){
    class2[[j]]<-ifelse(data>=cutpoint[j], "positive", "negative")
  }
  class<- data.frame(matrix(unlist(class2), nrow=uncertobj$singlefit$n, byrow=F),stringsAsFactors=TRUE)
  class$data<-data

  table<-vector("list", ncomp-1)
  for (i in 1:(ncomp-1)){
    table[[i]]<-table(class2[[i]])

  }
  npos<-vector("numeric")
  nneg<-vector("numeric")
  dpos<-vector("numeric")
  dneg<-vector("numeric")
  ppos<-vector("numeric")
  pneg<-vector("numeric")
  grpname<-vector("numeric")
  for (i in 1:(ncomp-1)){
    npos[i]<-table[[i]][2]
    nneg[i]<-table[[i]][1]
    dpos[i]<-npos[i]/(npos[i]+nneg[i])
    dneg[i]<-1-dpos[i]
    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
    grpname[i]<-paste("Cut between components ", i, " and ", i+1, sep="")
  }
  classtab<-data.frame( nneg, npos, dneg, dpos)
  outtab<-data.frame(nneg, npos, pneg, ppos)
  rownames(outtab) <- grpname
  colnames(outtab) <-c("Number Negative", "Number Positive", "Percent Negative", "Percent Positive")

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$g1<-g1
  cutobj$g2<-g2
  cutobj$classtab<-classtab
  cutobj$outtab<-outtab
  cutobj$cutnames<-grpname
  #    cutobj$uncertmat<-uncertobj$uncertmat
  #    cutobj$mcluncert<-uncertobj$mcluncert
  #    cutobj$restrmat<-uncertobj$restrmat
  #    cutobj$singlefit<-uncertobj$fit
  #    cutobj$data<-uncertobj$data
  #    cutobj$desc<-uncertobj$desc
  print(outtab)

  return(cutobj)
}

#-----------------------------------------------
#[IC7]  rawuncertgraph
#-----------------------------------------------
#Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){

  grobj<-list()
  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(uncertobj$datawithids$data)) 
    grobj$maxval<-max(uncertobj$datawithids$data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  plot(uncertobj$v, uncertobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", uncertobj$desc$dist, "with", uncertobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")

  col<-setcolor
  ncomp<-uncertobj$desc$ncomp
  for (i in 1:(uncertobj$desc$ncomp-1)){
    abline(v =multcutobj$cutpoint[i], untf = T, col=col[i], lwd=2)
  }

  if (suppresslegend==F) {
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 2:ncomp){
      legtxt[i]<-paste(multcutobj$cutnames[i-1], " , x=", round(multcutobj$cutpoint[i-1], 2), sep="")
      colors[i]<-col[i-1]
    }  
    legtxt[1]<-"Uncertainty"
    colors[1]<-col[length(col)]
    linewidth<-c(2)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=c(1,linewidth), legend=legtxt, xjust=1, seg.len=3)
  }
}

#-----------------------------------------------
#[IC8]  rawdistgraph
#-----------------------------------------------
#Displays the distributions of all the components for the chosen combination of distribution and number of components. 
rawdistgraph<-function(pickobj,multcutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp=pickobj$desc$ncomp
  grobj<-vector("list")
  grobj$gpar<-vector("list")
  data<-pickobj$datawithids$data
  color<-setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data)) 
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)

  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
  grobj$gt<-vector("list", ncomp+1)
  grobj$maxdens<-vector("numeric", ncomp)

  grobj$coloropt<-vector("list", 5)
  grobj$coloropt[[1]]<-c(color[1], color[9])
  grobj$coloropt[[2]]<-c(color[1], color[5], color[9])
  grobj$coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
  grobj$coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
  grobj$coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
  grobj$colors<-grobj$coloropt[[ncomp]]

  for (i in 1:ncomp){
    grobj$gpar$mean[i]=pickobj$par$mean[i]
    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
    grobj$gpar$shape[i]=pickobj$par$shape[i]
    grobj$gpar$pii[i]=pickobj$par$pii[i]
  }
  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
  for (i in 1:ncomp) {
    grobj$gt[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
  }
  grobj$gt[[ncomp+1]]<-vector("numeric", 1000)
  for (i in 1:ncomp) {
    grobj$gt[[ncomp+1]]<-grobj$gt[[ncomp+1]]+grobj$gt[[i]]
  }
  grobj$ylim<-c(0, max(grobj$gt[[ncomp+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
  grobj$xlim<-c(grobj$minval,grobj$maxval)

  dist<-c(0)[rep(c(0), times=ncomp)]
  for (i in 1:ncomp){
    if (pickobj$singlefit$shape[i]>0) {
      dist[i]<-1
    }
  }

  #Plot Graph

  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, pickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
  for (i in 1:ncomp) {
    lines(grobj$t, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
  }
  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
  legtxt<-vector("character", ncomp+1)
  colors<-vector("character", ncomp+1)
  for (i in 1:ncomp){
    legtxt[i]<-paste("Distribution of Component ", i, sep="")
    colors[i]<-grobj$colors[i]
  }  
  legtxt[ncomp+1]<-"Overall Distribution"
  colors[ncomp+1]<-grobj$colors[length(grobj$colors)]
  linetype<-c(1)[rep(c(1), times=ncomp)]
  if (suppresslegend==F) {
    legend("topright", col=colors, lwd=2, legend=legtxt, xjust=1, seg.len=3, title="Distributions", lty=c(linetype,2))
  }

  #return(grobj)
}

#-----------------------------------------------
#[IC9]  rawhistcuts
#-----------------------------------------------
#Displays which component each bar of the histogram would be assigned to.

rawhistcuts<-function(uncertobj,multcutobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  data<-uncertobj$datawithids$data
  grobj<-list()

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(data)) 
    grobj$maxval<-max(data)
    hist<-hist(data, breaks=setbreaks, plot=F)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
    datalim<-data[ which(data>=grobj$minval & data<=grobj$maxval)]
    hist<-hist(datalim, breaks=setbreaks, plot=F)
  }

  ncomp<-uncertobj$desc$ncomp
  cuts <- cut(hist$breaks, c(-Inf,multcutobj$cutpoint,Inf))

  color<-setcolor
  coloropt<-vector("list", 5)
  coloropt[[1]]<-c(color[1], color[9])
  coloropt[[2]]<-c(color[1], color[5], color[9])
  coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
  coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
  coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
  col<-coloropt[[ncomp]]

  cuts<-as.numeric(cuts)
  colors<-vector("character", length(cuts))
  for (i in 1:max(cuts)){
    for (j in 1:length(cuts)){
      if (cuts[j]==i) {
        colors[j]<-col[i]
      }
    }
  }
  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", uncertobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")

  if (suppresslegend==F){
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 1:ncomp){
      legtxt[i]<-paste("Component ", i, sep="")
      colors[i]<-col[i]
    }  
    linetype<-c(1)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=3, legend=legtxt, xjust=1, seg.len=3, title="Components", lty=linetype)
  }
}

#-----------------------------------------------
#[IC10] standindet
#-----------------------------------------------
#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (90 and 95), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
standindet<-function(uncertobj,multcutobj,cutcomp=0){
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp
  if (n==2){
    cutcomp=1
  } else if (cutcomp==0) {
    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
  }
  uncertdf<-multcutobj$uncertdf[[cutcomp]]
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  crosses<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    if (min.diffc<=0.001){
      y<-which(min.diffc==diff.cut)
      #take first one
      z<-min(y)
      cut<-uncertdf2$v[z]  
    } else {
      cut<-NA
    }
    return(cut)
  }
  cutpoint<-multcutobj$cutpoint[cutcomp]  
  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
  minuncub<-closest(uncertdf,cutpoint,max(data),0)
  lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
  ub90<-crosses(uncertdf,cutpoint,minuncub,0.1)
  lb95<-crosses(uncertdf,minunclb,cutpoint,0.05)  
  ub95<-crosses(uncertdf,cutpoint,minuncub,0.05)

  bound90<-c(lb90, ub90)
  bound95<-c(lb95, ub95)

  if (is.na(lb90)==T){
    message("Lower bound of the 90% indeterminate range not found because uncertainty is never below 10% to the left of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(ub90)==T){
    message("Upper bound of the 90% indeterminate range not found because uncertainty is never below 10% to the right of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(lb95)==T){
    message("Lower bound of the 95% indeterminate range not found because uncertainty is never below 5% to the left of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(ub95)==T){
    message("Upper bound of the 95% indeterminate range not found because uncertainty is never below 5% to the right of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }

  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
  class$group95<-"indeterminate"
  class$group95[class$data<=bound95[1]]<-"negative"
  class$group95[class$data>=bound95[2]]<-"positive"
  class$group90<-"indeterminate"
  class$group90[class$data<=bound90[1]]<-"negative"
  class$group90[class$data>=bound90[2]]<-"positive"

  class$groupdi<-as.factor(class$groupdi)
  class$group95<-as.factor(class$group95)
  class$group90<-as.factor(class$group90)

  class$group90 = factor(class$group90,levels(class$group90)[c(2,3,1)])
  class$group95 = factor(class$group95,levels(class$group95)[c(2,3,1)])

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$bound90<-bound90
  cutobj$bound95<-bound95
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$uncertobj<-uncertobj$uncertmat
  cutobj$density<-vector("list")
  cutobj$density$g1<-multcutobj$g1[[cutcomp]]
  cutobj$density$g2<-multcutobj$g2[[cutcomp]]
  cutobj$desc<-uncertobj$desc
  cutobj$desc$cutcomp<-cutcomp
  cutobj$datawithids<-uncertobj$datawithids

  return(cutobj) 
}

#-----------------------------------------------
#[IC11]  specindet
#-----------------------------------------------
#Provides cutpoint and allows the user to specify a maximum tolerable uncertianty which will be used to calculate an indeterminate ranges, as well as classification into negative, indeterminate and positive components. Maximum tolerable uncertianty is 1-certainty level (ex: uncertlevel=0.05 for 95% indetermintate range since the 95% certainty means that uncertainty of classification is less than 0.05 for all samples classified as positive or negative).
specindet<-function(uncertobj,multcutobj,certlevel,cutcomp=0){
  uncertlevel<-1-certlevel
  uncertlevelpct<-certlevel

  PD.v<-uncertobj$uncertmat
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp

  if (n==2){
    cutcomp=1
  } else if (cutcomp==0) {
    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
  }
  if (certlevel>=1 | certlevel<0.5){
    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 90% set certlevel=0.90")
  }

  uncertdf<-multcutobj$uncertdf[[cutcomp]]
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  crosses<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    if (min.diffc<=0.001){
      y<-which(min.diffc==diff.cut)
      #take first one
      z<-min(y)
      cut<-uncertdf2$v[z]  
    } else {
      cut<-NA
    }
    return(cut)
  }
  cutpoint<-closest(uncertdf,min(data),max(data),0.5)
  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
  minuncub<-closest(uncertdf,cutpoint,max(data),0)
  lbuns<-crosses(uncertdf,minunclb,cutpoint,uncertlevel)  
  ubuns<-crosses(uncertdf,cutpoint,minuncub,uncertlevel)
  bound<-c(lbuns, ubuns)

  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
  class$group<-"indeterminate"
  class$group[class$data<=bound[1]]<-"negative"
  class$group[class$data>=bound[2]]<-"positive"

  class$groupdi<-as.factor(class$groupdi)
  class$group<-as.factor(class$group)
  class$group = factor(class$group,levels(class$group)[c(2,3,1)])

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$bound<-bound
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Non.Standard"
  cutobj$uncertobj<-uncertobj$uncertmat
  cutobj$density<-data.frame(g1=multcutobj$g1[[cutcomp]], g2=multcutobj$g2[[cutcomp]])
  cutobj$desc<-uncertobj$desc
  cutobj$desc$cutcomp<-cutcomp
  cutobj$desc$uncertlevel<-uncertlevelpct
  cutobj$datawithids<-uncertobj$datawithids

  return(cutobj) 
}

#-----------------------------------------------
#[IC12]  cutuncertgraph
#-----------------------------------------------
#Displays the uncertainty function after components have been combined to create positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  color=setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty")
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty", xlim=c(grobj$minval,grobj$maxval))
  }

  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
  if (cutobj$type=="Standard") {
    abline(h =0.05, untf = T, col=color[4], lwd=1, lty=2)
    abline(h =0.1, untf = T, col=color[2], lwd=1, lty=2)
    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
    legtxt<-c("Uncertainty",
              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
              "Uncertainty=0.10",
              "Indeterminate Range with", 
              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
              "Uncertainty=0.05",
              "Indeterminate Range with", 
              paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
    )
    if (suppresslegend==F){
      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
      }
  } else {
    abline(h =(1-cutobj$desc$uncertlevel), untf = T, col=color[3], lwd=1, lty=2)
    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
    legtxt<-c("Uncertainty",
              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
              paste("Uncertainty=", (1-cutobj$desc$uncertlevel), sep =""),
              "Indeterminate Range with", 
              paste((100*cutobj$desc$uncertlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
    )
    if (suppresslegend==F){
      legend("topright", col=c(color[9],color[6], color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1), legend=legtxt, xjust=1, seg.len=1, title="Cutpoint, Uncertainties, and Boundaries")
    }
  }
}

#-----------------------------------------------
#[IC13]  cutdistgraph
#-----------------------------------------------
#Displays the distributions of the positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively). 
cutdistgraph<-function(cutobj,pickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp<-cutobj$desc$ncomp
  cutcomp<-cutobj$desc$cutcomp
  cutobj<-cutobj
  data<-pickobj$datawithids$data
  color<-setcolor

  grobj<-vector("list")
  grobj$gpar<-vector("list")

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) { 
    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data))     
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)

  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
  grobj$gs<-vector("list", ncomp+1)
  grobj$gt<-vector("list", 2+1)
  grobj$maxdens<-vector("numeric", ncomp)
  grobj$colors<-c(setcolor[1], setcolor[5], setcolor[9])
  grobj$t<-cutobj$uncertdf$v

  for (i in 1:ncomp){
    grobj$gpar$mean[i]=pickobj$par$mean[i]
    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
    grobj$gpar$shape[i]=pickobj$par$shape[i]
    grobj$gpar$pii[i]=pickobj$par$pii[i]
  }
  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
  for (i in 1:ncomp) {
    grobj$gs[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
  }
  grobj$gt[[1]]<-0
  grobj$gt[[2]]<-0
  grobj$gt[[3]]<-0
  for (i in 1:cutcomp){
    grobj$gt[[1]]<-grobj$gt[[1]]+grobj$gs[[i]]
  }
  for (i in (cutcomp+1):ncomp){
    grobj$gt[[2]]<-grobj$gt[[2]]+grobj$gs[[i]]
  }
  for (i in 1:ncomp) {
    grobj$gt[[3]]<-grobj$gt[[3]]+grobj$gs[[i]]
  }

  grobj$ylim<-c(0, max(grobj$gt[[2+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
  grobj$xlim<-c(round(min(pickobj$datawithids$data),digits=1),grobj$maxval)

  #Plot Graph
  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste(cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "), border=color[8], ylim=grobj$ylim, xlim=grobj$xlim)

  for (i in 1:2) {
    lines(cutobj$uncertdf$v, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
  }
  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
  if (cutobj$type=="Standard") {
    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
    if (suppresslegend==F) {
      legtxt<-c("Negative Distribution",
                "Positive Distribution",
                "Overall Distribution",
                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
                "Indeterminate Range with", 
                paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
                "Indeterminate Range with", 
                paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
      )
      legend("topright", col=c(color[1], color[5], color[9], color[6], color[2], "white", color[4], "white"), lwd=c(2,2,2,1,1,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1,1,1))
    }
  } else {
    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
    if (suppresslegend==F){
      legtxt<-c("Negative Distribution",
                "Positive Distribution",
                "Overall Distribution",
                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
                "Indeterminate Range with", 
                paste((cutobj$desc$uncertlevel*100), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
      )
      legend("topright", col=c(color[1], color[5], color[9], color[6], color[3], "white"), lwd=c(2,2,2,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1))
    }
  }

  #return(grobj)

}

#-----------------------------------------------
#[IC14]  outputdata
#-----------------------------------------------
outputdata<-function(standindetobj=NULL,specindetobj=NULL,fileandpathname=NULL){
  if (is.null(standindetobj)==T & is.null(specindetobj)==T) {
    print("At least one post-cut object must be included")
  } else {
    if (is.null(standindetobj)==T) {
      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
      dataout<-specindetobj$class
    } else if (is.null(specindetobj)==T){
      dataout<-standindetobj$class
    } else {
      datastand<-standindetobj$class[c("id", "group90", "group95")]
      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
      datanonst<-specindetobj$class
      dataout<-merge(datanonst,datastand,by.x="id", by.y="id")
    }
    if (is.null(fileandpathname)==F){
      write.csv(dataout, file =paste(fileandpathname, ".csv", sep=""), row.names=F)
    }
  }
  return(dataout)
}

#-----------------------------------------------
#[IC15]  summarytable
#-----------------------------------------------
#Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 90 and 95 indeterminate ranges and (if desired) one non-standard uncertainty level.
summarytable<-function(outdataobj,standindetobj=NULL,specindetobj=NULL){
  comptable <- function(table){
    dftable<-as.data.frame(table)
    len<-nrow(dftable)
    dftable$orders<-c(NA)[rep(c(1), times=len)]
    for (i in 1:len){
      if (as.character(dftable$Var1[i])=="negative"){
        dftable$orders[i]<-0
      } else if (as.character(dftable$Var1[i])=="indeterminate"){
        dftable$orders[i]<-1
      } else if (as.character(dftable$Var1[i])=="positive"){
        dftable$orders[i]<-2
      }
    }
    row.names(dftable)<-dftable$Var1
    dftable$Var1<-NULL
    if (is.na(dftable["indeterminate", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"indeterminate")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,1)
      dftable<-grps2
      Freq<-NULL
      x<-NULL
    }
    if (is.na(dftable["negative", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"negative")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,0)
      dftable<-grps2
      Freq<-NULL
      x<-NULL
    }
    if (is.na(dftable["positive", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"positive")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,2)
      dftable<-grps2    
      Freq<-NULL
      x<-NULL
    }
    dftable<-dftable[order(dftable$orders),]
    return(dftable)
  }

  grps<-table(outdataobj$groupdi)
  grps<-comptable(grps)
  grps$group<-row.names(grps)
  grps$groupdi<-grps$Freq
  grps$Freq<-NULL

  if (ncol(outdataobj)!=5 ){
    grpsns<-table(outdataobj[,4])
    grpsns<-comptable(grpsns)
    grps$gns<-grpsns$Freq
    names(grps)<-c(names(grps[1:3]),paste(names(outdataobj)[4]))
  }

  if (ncol(outdataobj)!=4 ){ 
    grps90<-table(outdataobj$group90)
    grps90<-comptable(grps90)
    grps$group90<-grps90$Freq
    grps95<-table(outdataobj$group95)
    grps95<-comptable(grps95)
    grps$group95<-grps95$Freq
  }

  names(grps)

  for (i in names(grps)){
    nneg<-vector("numeric")
    nind<-vector("numeric")
    npos<-vector("numeric")

    dneg<-vector("numeric")  
    dind<-vector("numeric")
    dpos<-vector("numeric")

    pneg<-vector("numeric")
    pind<-vector("numeric")
    ppos<-vector("numeric")
  }

  vars<-c("groupdi", "group90", "group95", "group80")

  grpname<-vector("numeric")
  for (i in 1:(ncol(grps)-2)){
    nneg[i]<-grps[1,i+2]
    nind[i]<-grps[2,i+2] 
    npos[i]<-grps[3,i+2]
  }
  for (i in 1:(ncol(grps)-2)){
    dneg[i]<-nneg[i]/(npos[i]+nind[i]+nneg[i])
    dind[i]<-nind[i]/(npos[i]+nind[i]+nneg[i])
    dpos[i]<-npos[i]/(npos[i]+nind[i]+nneg[i])
    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
    pind[i]<-paste(round(100*dind[i], 2), "%", sep="")
    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
  }
  if (ncol(grps)==4){
    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""))
  } else if (ncol(grps)==5) {
    grpname<-c("Raw Cutoff", "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
  } else if (ncol(grps)==6) {
    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""), "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
  }

  classtab<-data.frame(nneg, nind, npos, dneg, dind, dpos)
  outtab<-data.frame(nneg, nind, npos, pneg, pind, ppos)
  rownames(outtab) <- grpname
  colnames(outtab) <-c("No. Neg", "No. Indet", "No. Pos", "% Neg", "% Indet", "% Pos")

  model<-vector("list")
  if (is.null(specindetobj)==F){
    model$cutpoint<-specindetobj$cutpoint
    model$desc<-specindetobj$desc
    model$bounds<-vector("list")
    model$bounds$boundns<-specindetobj$bound
  } else {
    model$cutpoint<-standindetobj$cutpoint     
    model$desc<-standindetobj$desc
    model$bounds<-vector("list")
    model$bounds$bound90<-standindetobj$bound90
    model$bounds$bound95<-standindetobj$bound95
  }
  if (is.null(standindetobj)==F & is.null(specindetobj)==F){
    model$bounds$bound90<-standindetobj$bound90
    model$bounds$bound95<-standindetobj$bound95
  }
  model$classtab<-classtab
  model$outtab<-outtab
  model$classification<-outdataobj

  message(paste("The absolute cutoff is", round(standindetobj$cutpoint,3), sep=" "))
  message("Boundries of Indeterminates:")
  if (is.null(specindetobj)==F) {
    message(paste(specindetobj$desc$uncertlevel*100, "% Certainty Range: (", round(specindetobj$bound[1], 3), ", ", round(specindetobj$bound[2], 3), ")", sep=""))
  }
  message(paste("90% Certainty Range: (", round(standindetobj$bound90[1], 3), ", ", round(standindetobj$bound90[2], 3), ")", sep=""))
  message(paste("95% Certainty Range: (", round(standindetobj$bound95[1], 3), ", ", round(standindetobj$bound95[2], 3), ")", sep=""))

  return(model)
}

#-----------------------------------------------
#[IC16]  doallprobsfail
#-----------------------------------------------
#Performs all of the above functions, but is very unlikely to function since in the majority of cases a the two component case will not be optimal by BIC
doallprobsfail<-function(datawithids,fitres=NA){

  if (length(fitres)==1){
    fitres5<-fitloops(datawithids)
  }

  bestfit5<-bestfits(fitres5)
  bicgraph(bestfit5)
  pickobj5<-fitpick(fitres5,bestfit5)
  uncertobj5<-uncert(pickobj5)
  multcutobj5<-multcut(uncertobj5)
  rawdistgraph(pickobj5)
  rawuncertgraph(uncertobj5,multcutobj5)
  rawhistcuts(uncertobj5,multcutobj5)
  standindetpts5<-standindet(uncertobj5,multcutobj5)
  cutuncertgraph(standindetpts5)
  cutdistgraph(standindetpts5, pickobj5)
  nsindet5<-nsstandindets(uncertobj5,multcutobj5,certlevel=0.8,cutcomp=1)
  cutuncertgraph(nsindet5)
  cutdistgraph(nsindet5, pickobj5)
  outdata5<-outputdata(standindetpts5,nsindet5)
  summary<-summarytable(outdata5, standindetpts5, nsindet5)

}

#-----------------------------------------------
#[IC17]  twobestdorest
#-----------------------------------------------
#Performs all functions after bestfit if the optimal number of components is 2.
twobestdorest<-function(fitloopsobj,bestfit){

  pickobj<-fitpick(fitloopsobj,bestfit)
  uncertobj<-uncert(pickobj)
  multcutobj<-multcut(uncertobj)
  rawdistgraph(pickobj)
  rawuncertgraph(uncertobj,multcutobj)
  rawhistcuts(uncertobj,multcutobj)
  scptsobj<-standindet(uncertobj,multcutobj)
  cutuncertgraph(scptsobj)
  cutdistgraph(scptsobj, pickobj)
  outdata<-outputdata(standindetpts,nsindet)
  summary<-summarytable(outdata, standindetpts, nsindet)

  model<-vector("list")
  model$uncertobj<-uncertobj
  model$multcutobj<-multcutobj
  model$scptsobj<-scptsobj
  model$outdataobj<-outdata
  model$summaryobj<-summary

  return(model)

}

```

```{r example1}

examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv")

hist(examp1$data, breaks=100, main="Histogram of Data")
dens<-density(examp1$data)
plot(dens, main="Density of Data")

fitloopsobject<-fitloops(datawithids=examp1)
bestfitsobject<-bestfits(fitres=fitloopsobject)
bicgraph(model=bestfitsobject)

fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)

uncertobject<-uncert(fitpickobj=fitpickobject)
multcutobject<-multcut(uncertobj=uncertobject)

rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
rawdistgraph(pickobj=fitpickobject)
rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)

standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)

cutuncertgraph(cutobj=standindetobject)
cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
cutuncertgraph(cutobj=specindetobject)
cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
outtable<-summaryobject$outtab
outtable

```

```{r example2}

examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", colClasses=c("NULL", NA, NA))

hist(examp2$data, breaks=100, main="Histogram of Data")
dens<-density(examp2$data)
plot(dens, main="Density of Data")

fitloopsobject<-fitloops(datawithids=examp2)
#might not run, if it doesn't, then load the full file: 
load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2fitloops")
fitloopsobject<-ex2
bestfitsobject<-bestfits(fitres=fitloopsobject)
bicgraph(model=bestfitsobject)

fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)

uncertobject<-uncert(fitpickobj=fitpickobject)
multcutobject<-multcut(uncertobj=uncertobject)

rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
rawdistgraph(pickobj=fitpickobject)
rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)

standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)

cutuncertgraph(cutobj=standindetobject)
cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
cutuncertgraph(cutobj=specindetobject)
cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
outtable<-summaryobject$outtab
outtable

```

```{r while loop}

stop<-F
iter<-0
comp<-0

  loopErr = function(data) {
    tryCatch(smsn.mix(data, nu =3, g = 5, get.init = TRUE, criteria = TRUE, group = TRUE, family = "Normal", calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 100, n.start = 50)),
             error = function(e) {message(paste("Error in iteration", iter, sep=" ")); 
               NA}) 
  }

#fit5<-list()
#while (stop==F){
#  iter<-iter+1
#  set.seed(iter*10000+iter^2+iter)    
#  fit5[iter]<- loopErr(examp2$data)
#  if (is.na(fit5[iter])==F){
#    stop<-T
#  }
#}

```

### 2017-03-09 PaperEx170308.Rmd
results for paper

@@
 
-```{r loadfunctions}
+```{r set_knitr_chunk_options, echo = F, warning=F, message=F}
+rm(list = ls())
+
+library(knitr)
+library(markdown)
+opts_chunk$set(message=FALSE, echo = F,
+               warning = F,fig.height = 5, fig.width = 10, tidy = F)
+
+#Work
+#opts_knit$set(root.dir = "C:/Users/ntdintern/Google Drive/Sarah Sullivan/Code")
+#School
+#opts_knit$set(root.dir = "H:/TFGH/Code")
+#Laptop
+opts_knit$set(root.dir = "/Users/sarahsullivan/Google Drive/Sarah Sullivan/Code")
+```
+
+```{r loadfunctions, echo=F, warning=F, message=F}
 library(mixsmsn)
@@
 
-fitloopsobject<-fitloops(datawithids=examp2)
-#might not run, if it doesn't, then load the full file: 
+#fitloopsobject<-fitloops(datawithids=examp2)
+  #might not run due to 5 component solution not being easily optimized, if it doesn't run, either load the full file, or run fitloopsobject<-fitloops(datawithids=examp2,maxcp=4): 
 load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2fitloops")
 fitloopsobject<-ex2
+
 bestfitsobject<-bestfits(fitres=fitloopsobject)
@@
 
-```{r while loop}
+```{r while loop, echo=F}
 

### 2017-03-10 Cutoff_Algorithm.R
consistently "skew-normal" not skewnormal or skew normal

@@ -90,7 +90,7 @@ library("sn")
 #[IC1]  fitloops
 #-----------------------------------------------
 
-#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skewnormal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
+#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skew-normal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
 #accepts a dataframe with columns id and data
 fitloops<-function(datawithids,loops=10,maxcp=5){
   dists="Both"
@@ -115,13 +115,13 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       norms$ms[[i]][[j]]<-vector("list", 15)
       if (dists=="Both") {
         norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")  
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
+        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew-normal")
       } else if (dists=="Normal"){ 
         norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")
-      } else if (dists=="Skew.normal"){
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
+      } else if (dists=="Skew-normal"){
+        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew-normal")
       } else {
[…]

### 2017-03-10 Cutoff_Algorithm.R
fix "skew-normal" for algorithm (Skew.normal)

@@ -115,11 +115,11 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       norms$ms[[i]][[j]]<-vector("list", 15)
       if (dists=="Both") {
         norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")  
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew-normal")
+        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
       } else if (dists=="Normal"){ 
         norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")
-      } else if (dists=="Skew-normal"){
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew-normal")
+      } else if (dists=="Skew.normal"){
+        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
       } else {
         stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew-normal'"))
       }
@@ -138,7 +138,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       if (dists=="Normal"){ 
         norms$ms[[i]][[j]]<-vector("list")
         norms$ms[[i]][[j]]$bic<-NA
-      } else if (dists=="Skew-normal"){
+      } else if (dists=="Skew.normal"){
         norms$mc[[i]][[j]]<-vector("list")
         norms$mc[[i]][[j]]$bic<-NA
       }
@@ -165,7 +165,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       } else {
         nobj$lmin$mc[i]<-1
       }
-      if (dists=="Both" | dists=="Skew-normal"){
+      if (dists=="Both" | dists=="Skew.normal"){
         nobj$lmin$ms[i]<-which.min(nobj$ms[i,])
       } else {
         nobj$lmin$ms[i]<-1
@@ -188,12 +188,12 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
     if (which.max(errors)<=maxcp){
       dist<-"Normal"
     } else {
-      dist<-"Skew-normal"
+      dist<-"Skew.normal"
     }
     if (ic<=2){
[…]

### 2017-03-10 Cutoff_Algorithm.R
edit ylim in bic graph

@@ -301,11 +301,13 @@ bestfits<-function(fitres){
 
 bicgraph<-function(model,title="BIC by type and number of distributions",setcolor=c("magenta", "blue", "black")){
   color<-setcolor
+  buffer<-(max(model$bictab)-min(model$bictab))/20
+  ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
   if (is.na(model$bictab$Normal[1])==F) {
-    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n")
+    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
     lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
   } else {
-    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n")
+    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n", ylim=ylim)
     lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
   }  
   points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)

### 2017-03-11 PaperEx170308.Rmd
EOD 170310 -Both algorithm and paper are in good shape -Upcoming Algorithm tasks: Overhaul to 4 key functions 1) Fitting: fitloops and bestfit    -BIC plot 2) Selection: fitpick uncert and multcut    -rawdist    -rawuncert    -rawhist 3) Cutoff: standard and non-standard cuts    -cutuncert    -cutdist 4) Summarization/Exportation: outputdata and summarytable

@@
 
-```{r loadfunctions, echo=F, warning=F, message=F}
-library(mixsmsn)
-library(sn)
+```{r loadfunctions, echo=F}
+
+#-----------------------------------------------
+#[IB] Load Packages
+#-----------------------------------------------
+
+library("mixsmsn")
+library("sn")
+
 #-----------------------------------------------
@@
 This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails.
-The default values are to do both skewnormal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components.
+The default values are to do both skew-normal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components.
 It is not reccommended to change these unless the default function fails.
@@
       } else {
-        stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew.normal'"))
+        stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew-normal'"))
       }
@@
     if (ic<=2){
-      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s).
+      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
 First ensure that the necessary packages have been loaded correctly.
-If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skewnormal or normal distribution assumptions."))
-    } else {
-      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s).
+If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
+    } else {
+      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
 Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
@@
   } else if (nobj$dist=="sn") {
-    nobj$distp="Skew normal"
-    nobj$distf="Skew.normal"
+    nobj$distp="Skew-normal"
+    nobj$distf="Skew-normal"
   } else {
@@
   rownames(model$bictab)<-componentnames
-  colnames(model$bictab)<-c("Normal", "Skew normal")
-  model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
-  rownames(model$summary)<-c("Best Overall", "Best Skew normal  ", "Best Normal", "Two Skew normal", "Two Normal")
+  colnames(model$bictab)<-c("Normal", "Skew-normal")
+  model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew-normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew-normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
+  rownames(model$summary)<-c("Best Overall", "Best Skew-normal  ", "Best Normal", "Two Skew-normal", "Two Normal")
   colnames(model$summary)<-c("Description         ", "BIC  ")
@@
 
-bicgraph<-function(model,title="BIC by type and number of distributions",setcolor=c("magenta", "blue", "black")){
+bicgraph<-function(model,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
   color<-setcolor
+  buffer<-(max(model$bictab)-min(model$bictab))/20
+  ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
   if (is.na(model$bictab$Normal[1])==F) {
-    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n")
-    lines(model$bictab$`Skew normal`, type="o", pch=2, lty=2, col=color[2])
-  } else {
-    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n")
-    lines(model$bictab$`Skew normal`, pch=1, col=color[1], type="o")    
+    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of components", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
+    lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
+  } else {
+    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of components", ylab="BIC",  xaxt="n", ylim=ylim)
+    lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
   }  
@@
   axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
-  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))
+  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew-normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))
 
@@
   n=ncomp
-  if (dist!="Normal" & dist!="Skew.normal") {
-    stop("Distribution must be specified as either 'Normal' or 'Skew.normal'")
-  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew.normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
+  if (dist!="Normal" & dist!="Skew-normal") {
+    stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
+  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
     stop("Must choose a distribution and number of components that were fitted in the fitloops function")
@@
     } else if (dist=="sn") {
-      distp="Skew Normal"
+      distp="Skew-Normal"
     } else {
@@
 }
-
 ```
 
-```{r example1}
-
-examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv")
+#Example 1
+```{r example1, echo=T}
+
+examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))
 
 hist(examp1$data, breaks=100, main="Histogram of Data")
-dens<-density(examp1$data)
-plot(dens, main="Density of Data")
-
-fitloopsobject<-fitloops(datawithids=examp1)
-bestfitsobject<-bestfits(fitres=fitloopsobject)
-bicgraph(model=bestfitsobject)
-
-fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)
-
-uncertobject<-uncert(fitpickobj=fitpickobject)
-multcutobject<-multcut(uncertobj=uncertobject)
-
-rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
-rawdistgraph(pickobj=fitpickobject)
-rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)
-
-standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
-specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)
-
-cutuncertgraph(cutobj=standindetobject)
-cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
-cutuncertgraph(cutobj=specindetobject)
-cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
-outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
-summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
-outtable<-summaryobject$outtab
-outtable
+dens1<-density(examp1$data)
+plot(dens1, main="Density of Data")
+
+#fitloopsobject<-fitloops(datawithids=examp2)
+  #might not run due to 5 component solution not being easily optimized, if it doesn't run either load the full file, or run fitloopsobject<-fitloops(datawithids=examp2,maxcp=4): 
+load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")
+
+fitloopsobject1<-ex1
+
+bestfitsobject1<-bestfits(fitres=fitloopsobject1)
+bicgraph(model=bestfitsobject1)
+
+fitpickobject1<-fitpick(fitres=fitloopsobject1,bestfits=bestfitsobject1)
+
+uncertobject1<-uncert(fitpickobj=fitpickobject1)
+multcutobject1<-multcut(uncertobj=uncertobject1)
+
+rawuncertgraph(uncertobj=uncertobject1,multcutobj=multcutobject1)
+rawdistgraph(pickobj=fitpickobject1)
+rawhistcuts(uncertobj=uncertobject1,multcutobj=multcutobject1)
+
+standindetobject1<-standindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1)
+specindetobject1<-specindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1,certlevel=0.8)
+
+cutuncertgraph(cutobj=standindetobject1)
+cutdistgraph(cutobj=standindetobject1, pickobj=fitpickobject1)
+cutuncertgraph(cutobj=specindetobject1)
+cutdistgraph(cutobj=specindetobject1, pickobj=fitpickobject1)
+outdataobject1<-outputdata(standindet=standindetobject1,specindet=specindetobject1)
+summaryobject1<-summarytable(outdataobj=outdataobject1, standindetobj=standindetobject1, specindetobj=specindetobject1)
+outtable1<-summaryobject1$outtab
+outtable1
 
@@
 
-```{r example2}
-
-examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", colClasses=c("NULL", NA, NA))
+#Example 2
+
+```{r example2, echo=T}
+
+examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv")
 
 hist(examp2$data, breaks=100, main="Histogram of Data")
-dens<-density(examp2$data)
-plot(dens, main="Density of Data")
-
-#fitloopsobject<-fitloops(datawithids=examp2)
-  #might not run due to 5 component solution not being easily optimized, if it doesn't run, either load the full file, or run fitloopsobject<-fitloops(datawithids=examp2,maxcp=4): 
-load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2fitloops")
-fitloopsobject<-ex2
-
-bestfitsobject<-bestfits(fitres=fitloopsobject)
-bicgraph(model=bestfitsobject)
-
-fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)
-
-uncertobject<-uncert(fitpickobj=fitpickobject)
-multcutobject<-multcut(uncertobj=uncertobject)
-
-rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
-rawdistgraph(pickobj=fitpickobject)
-rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)
-
-standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
-specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)
-
-cutuncertgraph(cutobj=standindetobject)
-cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
-cutuncertgraph(cutobj=specindetobject)
-cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
-outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
-summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
-outtable<-summaryobject$outtab
-outtable
+dens2<-density(examp2$data)
+plot(dens2, main="Density of Data")
+
+fitloopsobject2<-fitloops(datawithids=examp2)
+bestfitsobject2<-bestfits(fitres=fitloopsobject2)
+bicgraph(model=bestfitsobject2)
+
+fitpickobject2<-fitpick(fitres=fitloopsobject2,bestfits=bestfitsobject2)
+
+uncertobject2<-uncert(fitpickobj=fitpickobject2)
+multcutobject2<-multcut(uncertobj=uncertobject2)
+
+rawuncertgraph(uncertobj=uncertobject2,multcutobj=multcutobject2)
+rawdistgraph(pickobj=fitpickobject2)
+rawhistcuts(uncertobj=uncertobject2,multcutobj=multcutobject2)
+
+standindetobject2<-standindet(uncertobj=uncertobject2,multcutobj=multcutobject2,cutcomp=1)
+specindetobject2<-specindet(uncertobj=uncertobject2,multcutobj=multcutobject2,cutcomp=1,certlevel=0.8)
+
+cutuncertgraph(cutobj=standindetobject2, xlim=c(min(examp2$data),3))
+cutdistgraph(cutobj=standindetobject2, pickobj=fitpickobject2, xlim=c(min(examp2$data),3))
+cutuncertgraph(cutobj=specindetobject2, xlim=c(min(examp2$data),3))
+cutdistgraph(cutobj=specindetobject2, pickobj=fitpickobject2, xlim=c(min(examp2$data),3))
+outdataobject2<-outputdata(standindet=standindetobject2,specindet=specindetobject2)
+summaryobject2<-summarytable(outdataobj=outdataobject2, standindetobj=standindetobject2, specindetobj=specindetobject2)
+outtable2<-summaryobject2$outtab
+outtable2
 
 ```
-
-```{r while loop, echo=F}
-
-stop<-F
-iter<-0
-comp<-0
-
-  loopErr = function(data) {
-    tryCatch(smsn.mix(data, nu =3, g = 5, get.init = TRUE, criteria = TRUE, group = TRUE, family = "Normal", calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 100, n.start = 50)),
-             error = function(e) {message(paste("Error in iteration", iter, sep=" ")); 
-               NA}) 
-  }
-
-#fit5<-list()
-#while (stop==F){
-#  iter<-iter+1
-#  set.seed(iter*10000+iter^2+iter)    
-#  fit5[iter]<- loopErr(examp2$data)
-#  if (is.na(fit5[iter])==F){
-#    stop<-T
-#  }
-#}
-
-```

### 2017-03-11 Cutoff_Algorithm.R

@@ -299,15 +299,15 @@ bestfits<-function(fitres){
 #-----------------------------------------------
 #Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.
 
-bicgraph<-function(model,title="BIC by type and number of distributions",setcolor=c("magenta", "blue", "black")){
+bicgraph<-function(model,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
   color<-setcolor
   buffer<-(max(model$bictab)-min(model$bictab))/20
   ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
   if (is.na(model$bictab$Normal[1])==F) {
-    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
+    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of components", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
     lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
   } else {
-    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n", ylim=ylim)
+    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of components", ylab="BIC",  xaxt="n", ylim=ylim)
     lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
   }  
   points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
@@ -973,6 +973,7 @@ cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppressleg
   "black"            #9
 )){
   color=setcolor
+  grobj<-vector("list")
   
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
[…]

### 2017-03-19 PaperEx170308.Rmd
Changed from 90&95 to 80&90

@@
 standindetobject1<-standindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1)
-specindetobject1<-specindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1,certlevel=0.8)
+specindetobject1<-specindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1,certlevel=0.95)
 

### 2017-03-19 Cutoff_Algorithm.R

@@ -787,7 +787,7 @@ rawhistcuts<-function(uncertobj,multcutobj,xlab="Optical Density",xlim=c(NA,NA),
 #-----------------------------------------------
 #[IC10] standindet
 #-----------------------------------------------
-#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (90 and 95), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
+#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (80 and 90), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
 standindet<-function(uncertobj,multcutobj,cutcomp=0){
   data<-uncertobj$datawithids$data
   n=uncertobj$desc$ncomp
@@ -824,47 +824,47 @@ standindet<-function(uncertobj,multcutobj,cutcomp=0){
   cutpoint<-multcutobj$cutpoint[cutcomp]  
   minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
   minuncub<-closest(uncertdf,cutpoint,max(data),0)
+  lb80<-crosses(uncertdf,minunclb,cutpoint,0.2)  
+  ub80<-crosses(uncertdf,cutpoint,minuncub,0.2)
   lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
   ub90<-crosses(uncertdf,cutpoint,minuncub,0.1)
-  lb95<-crosses(uncertdf,minunclb,cutpoint,0.05)  
-  ub95<-crosses(uncertdf,cutpoint,minuncub,0.05)
   
+  bound80<-c(lb80, ub80)
[…]

### 2017-03-19 Cutoff_Algorithm.R
Start of condensing functions, fitloops and bestfit have not yet been successfully merged

@@ -183,7 +183,88 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
     fitres$maxcp<-maxcp
     fitres$datawithids<-datawithids
     fitres$errors<-errors
-    return(fitres)
+  
+    nobj<-vector("list")
+    
+    for (i in 1:maxcp){
+      for (j in 1:loops){
+        nobj$no[i]<-fitres$no[[i]]$bic
+        nobj$sn[i]<-fitres$sn[[i]]$bic
+      }
+    }
+    
+    nobj$minind<-data.frame(no=as.numeric(c(NA)), sn=as.numeric(c(NA)))
+    
+    
+    
+    if (is.na(nobj$no[1])==T){
+      nobj$minind$no<-1
+    } else {
+      nobj$minind$no <- (which.min(nobj$no))
+    }
+    
+    if (is.na(nobj$sn[1])==T){
+      nobj$minind$sn<-1
+    } else {
+      nobj$minind$sn <- (which.min(nobj$sn))
+    }
+    
+    nobj$min<-data.frame(type=c("no", "sn"), bic=as.numeric(c(NA, NA)))
+    
+    
+    nobj$min$type<-as.character(c("no", "sn"))
+    type<-c("no", "sn")
+    nobj$min$type<-type
+    
+    nobj$min[1,2]<-fitres$no[[nobj$minind$no[1]]]$bic
+    nobj$min[2,2]<-fitres$sn[[nobj$minind$sn[1]]]$bic
+    
+    nobj$distnum<-which.min(nobj$min$bic)
+    nobj$dist<-nobj$min$type[nobj$distnum]
+    
+    nobj$best<-fitres[[nobj$dist]][[nobj$minind[nobj$dist][1,1]]]
+    
+    if (nobj$dist=="no"){
+      nobj$distp="Normal"
+      nobj$distf="Normal"
+    } else if (nobj$dist=="sn") {
+      nobj$distp="Skew-normal"
+      nobj$distf="Skew-normal"
+    } else {
+      nobj$distp=""
+    }
+    
+    model<-vector("list")
+    model$sn2<-fitres$sn[[2]]
+    model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
+    model$norm2<-fitres$no[[2]]
+    model$normbest<-fitres$no[[nobj$minind$no[1]]]
+    model$best<-nobj$best
[…]

### 2017-03-19 Cutoff_Algorithm.R
successful combination of fitloops and bestfit

@@ -100,7 +100,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
   norms<-vector("list")
   errors<-matrix(0, nrow=maxcp, ncol=2)
   loopErr = function(data,i,j,family) {
-    tryCatch(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5)),
+    tryCatch(suppressWarnings(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5))),
              error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
                NA}) 
   }
@@ -179,10 +179,6 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
       fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
     }
-    fitres$loops<-loops
-    fitres$maxcp<-maxcp
-    fitres$datawithids<-datawithids
-    fitres$errors<-errors
   
     nobj<-vector("list")
     
@@ -235,6 +231,13 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
     }
     
     model<-vector("list")
+    
+    model$loops<-loops
+    model$maxcp<-maxcp
+    model$datawithids<-datawithids
+    model$errors<-errors
+    model$fitres<-fitres
+    
     model$sn2<-fitres$sn[[2]]
     model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
     model$norm2<-fitres$no[[2]]
@@ -253,10 +256,9 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
     rownames(model$summary)<-c("Best Overall", "Best Skew-normal  ", "Best Normal", "Two Skew-normal", "Two Normal")
     colnames(model$summary)<-c("Description         ", "BIC  ")
     
-    
     model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
-    model$datawithids<-fitres$datawithids
     
+    print("")
[…]

### 2017-03-20 Cutoff_Algorithm.R
start to combine fitpick uncert and multcut

@@ -312,265 +312,291 @@ bicgraph<-function(model,title="BIC by type and number of components",setcolor=c
 #-----------------------------------------------
 #[IC4]  fitpick
 #-----------------------------------------------
-#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
 
-fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
-  if (dist=="" & is.na(ncomp)==T) {
-    dist=bestfits$bestdesc$dist
-    ncomp=bestfits$bestdesc$ncomp
-  }
-  if (ncomp==1) {
-    stop("Single component specified so there is no cut-point")
+cutfn<-function(fitresobj,dist="",ncomp=NA){
+  #Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
+  fitpick<-function(fitres,dist="",ncomp=NA){
+    if (dist=="" & is.na(ncomp)==T) {
+      dist=bestfits$bestdesc$dist
+      ncomp=bestfits$bestdesc$ncomp
+    }
+    if (ncomp==1) {
+      stop("Single component specified so there is no cut-point")
+    }
+    
+    data=fitres$datawithids$data
+    n=ncomp
+    if (dist!="Normal" & dist!="Skew-normal") {
+      stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
+    } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
+      stop("Must choose a distribution and number of components that were fitted in the fitloops function")
+    } else {
+      if (dist=="Normal"){
+        dist="no"
+      } else {
+        dist="sn"
+      }
+      
+      if (dist=="no"){
+        distp="Normal"
+      } else if (dist=="sn") {
+        distp="Skew-Normal"
+      } else {
+        distp=""
+      }
+      
[…]

### 2017-03-20 Cutoff_Algorithm.R
Cutpick construction: -fitpick function updated toward cutpick

@@ -236,7 +236,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
     model$maxcp<-maxcp
     model$datawithids<-datawithids
     model$errors<-errors
-    model$fitres<-fitres
+    model$loopfits<-fitres
     
     model$sn2<-fitres$sn[[2]]
     model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
@@ -286,7 +286,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
   
 
 #-----------------------------------------------
-#[IC3]  bicgraph
+#[IC2]  bicgraph
 #-----------------------------------------------
 #Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.
 
@@ -310,7 +310,7 @@ bicgraph<-function(model,title="BIC by type and number of components",setcolor=c
 
 
 #-----------------------------------------------
-#[IC4]  fitpick
+#[IC3]  cutfn
 #-----------------------------------------------
 
 cutfn<-function(fitresobj,dist="",ncomp=NA){
@@ -595,6 +595,270 @@ cutfn<-function(fitresobj,dist="",ncomp=NA){
   
   return()
   
+}
+
+
+#-----------------------------------------------
+#[IC4]  fitpick
+#-----------------------------------------------
+#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
+
+cutpick<-function(fitres, dist="", ncomp=NA){
+  #fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
+  if (dist=="" & is.na(ncomp)==T) {
+    dist=fitres$bestdesc$dist
+    ncomp=fitres$bestdesc$ncomp
+  }
[…]

### 2017-03-20 Cutoff_Algorithm.R
Finished writing the cutpick function

@@ -307,301 +307,10 @@ bicgraph<-function(model,title="BIC by type and number of components",setcolor=c
   
 }
 
-
-
 #-----------------------------------------------
-#[IC3]  cutfn
+#[IC3]  cutpick
 #-----------------------------------------------
-
-cutfn<-function(fitresobj,dist="",ncomp=NA){
-  #Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
-  fitpick<-function(fitres,dist="",ncomp=NA){
-    if (dist=="" & is.na(ncomp)==T) {
-      dist=bestfits$bestdesc$dist
-      ncomp=bestfits$bestdesc$ncomp
-    }
-    if (ncomp==1) {
-      stop("Single component specified so there is no cut-point")
-    }
-    
-    data=fitres$datawithids$data
-    n=ncomp
-    if (dist!="Normal" & dist!="Skew-normal") {
-      stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
-    } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
-      stop("Must choose a distribution and number of components that were fitted in the fitloops function")
-    } else {
-      if (dist=="Normal"){
-        dist="no"
-      } else {
-        dist="sn"
-      }
-      
-      if (dist=="no"){
-        distp="Normal"
-      } else if (dist=="sn") {
-        distp="Skew-Normal"
-      } else {
-        distp=""
-      }
-      
-      singlefit<-vector("list")
-      singlefit<-fitres[[dist]][[ncomp]]  
-      
-      comp<-data.frame(ogroup=singlefit$group, data=data)
-      comp$index <- as.numeric(row.names(comp))
-      comp <- comp[order(comp$data),] 
-      comp$diff<-c(0, diff(comp$ogroup))
-      nl=as.numeric(dim(comp)[1])
-      
-      #Will recode the distributions so that they are in order with 1 being the lowest
[…]

### 2017-03-20 Cutoff_Algorithm.R
Update later functions to work with cutpick

@@ -545,7 +545,7 @@ cutpick<-function(fitres, dist="", ncomp=NA){
 #[IC7]  rawuncertgraph
 #-----------------------------------------------
 #Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+rawuncertgraph<-function(cutpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
   "green4",          #1
   "turquoise3",      #2
   "royalblue2",      #3
@@ -559,26 +559,26 @@ rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppr
   
   grobj<-list()
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(uncertobj$datawithids$data)) 
-    grobj$maxval<-max(uncertobj$datawithids$data)
+    grobj$minval<-(1/10)*trunc(10*min(cutpickobj$datawithids$data)) 
+    grobj$maxval<-max(cutpickobj$datawithids$data)
   } else {
     grobj$minval<-xlim[1]
     grobj$maxval<-xlim[2]
   }
 
-  plot(uncertobj$v, uncertobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", uncertobj$desc$dist, "with", uncertobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
+  plot(cutpickobj$v, cutpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", cutpickobj$desc$dist, "with", cutpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
     
[…]

### 2017-03-20 Cutoff_Algorithm.R
EOD 170319 Finished: -fitloops -cutpick -bicgraph -rawuncertgraph -rawdistgraph -rawhistcuts Need to fix: -indeterminate (standard and nonstandard) -cutuncertgraph -cutdistgraph -outputdata -summarytable -twobestdorest

@@ -1292,7 +1292,7 @@ doallprobsfail<-function(datawithids,fitres=NA){
   summary<-summarytable(outdata5, standindetpts5, nsindet5)
   
 }
-
+ 
 #-----------------------------------------------
 #[IC17]  twobestdorest
 #-----------------------------------------------

### 2017-03-21 Cutoff_Algorithm.R
rename cutpick to modelpick and update following functions

@@ -308,11 +308,11 @@ bicgraph<-function(model,title="BIC by type and number of components",setcolor=c
 }
 
 #-----------------------------------------------
-#[IC3]  cutpick
+#[IC3]  modelpick
 #-----------------------------------------------
 #Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise. The function also performs uncertainty calculations and calculates raw percentages and counts positive and negative for all possible cut points between distributions
 
-cutpick<-function(fitres, dist="", ncomp=NA){
+modelpick<-function(fitres, dist="", ncomp=NA){
   #fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
   if (dist=="" & is.na(ncomp)==T) {
     dist=fitres$bestdesc$dist
@@ -542,10 +542,10 @@ cutpick<-function(fitres, dist="", ncomp=NA){
 }
   
 #-----------------------------------------------
-#[IC7]  rawuncertgraph
+#[IC4]  rawuncertgraph
 #-----------------------------------------------
 #Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(cutpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
   "green4",          #1
   "turquoise3",      #2
   "royalblue2",      #3
@@ -559,26 +559,26 @@ rawuncertgraph<-function(cutpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=
   
   grobj<-list()
[…]

### 2017-03-21 Cutoff_Algorithm.R
Cutoff function created and internal calculations updated. Still need logic to create new dtructure for standard and non-standard and description etc

@@ -746,16 +746,22 @@ rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreak
 }
 
 #-----------------------------------------------
-#[IC7] standindet
+#[IC7] cutoff
 #-----------------------------------------------
 #After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (80 and 90), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
-standindet<-function(modelpickobj,cutcomp=0){
+cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
   data<-modelpickobj$datawithids$data
   n=modelpickobj$desc$ncomp
   if (n==2){
     cutcomp=1
   } else if (cutcomp==0) {
     stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
+  }
+  if ((certlevel>=1 & certlevel!=0) | (certlevel<0.5 & certlevel!=0)){
+    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 80% set certlevel=0.80")
+  }
+  if (standcert==F & newcertlevel==0){
+    stop("Neither standard nor non-standard certainty levels are requested. For standard certainty levels set standardcert=T. For non-standard certainty levels set newcertlevel to desired certainty. Both standard and nonstandard certainties can be generated in the same function call.")
   }
   uncertdf<-modelpickobj$uncertdf[[cutcomp]]
   closest<-function(uncertdf,min,max,level){
@@ -785,26 +791,34 @@ standindet<-function(modelpickobj,cutcomp=0){
   cutpoint<-modelpickobj$cutpoint[cutcomp]  
   minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
   minuncub<-closest(uncertdf,cutpoint,max(data),0)
-  lb80<-crosses(uncertdf,minunclb,cutpoint,0.2)  
-  ub80<-crosses(uncertdf,cutpoint,minuncub,0.2)
-  lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
[…]

### 2017-03-21 Cutoff_Algorithm.R
EOD 170321 Correct but need to be tested: -fitloops -bicgraph -modelpick -rawuncertgraph -rawdistgraph -rawhistcuts -cutoff Not updated and not compatible: -cutuncertgraph: need to adjust for "Standard.and.Non" -cutdistgraph: need to adjust for "Standard.and.Non" -summaryout: need to combine summarytable and outoutdata

@@ -24,22 +24,15 @@
 #[IB] Load Packages
 #[IC] Load Functions
 #[IC1]  fitloops
-#[IC2]  bestfits
-#[IC3]  bicgraph
-#[IC4]  fitpick
-#[IC5]  uncert
-#[IC6]  multcut
-#[IC7]  rawuncertgraph
-#[IC8]  rawdistgraph
-#[IC9]  rawhistcuts
-#[IC10] standindet
-#[IC11] specindet
+#[IC2]  bicgraph
+#[IC3]  modelpick
+#[IC4]  rawuncertgraph
+#[IC5]  rawdistgraph
+#[IC6]  rawhistcuts
+#[IC7]  cutoff
 #[IC12] cutuncertgraph
 #[IC13] cutdistgraph
-#[IC14] outputdata
-#[IC15] summarytable
-#[IC16] doallprobsfail
-#[IC17] twobestdorest
+#[IC14] summaryout
 #-----------------------------------------------
 #[II] Load and clean data
 #[IIA] Import data 
@@ -746,12 +739,13 @@ rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreak
 }
 
 #-----------------------------------------------
-#[IC7] cutoff
+#[IC7]  cutoff
 #-----------------------------------------------
 #After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (80 and 90), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
 cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
   data<-modelpickobj$datawithids$data
   n=modelpickobj$desc$ncomp
+  certlevel=newcertlevel
   if (n==2){
     cutcomp=1
   } else if (cutcomp==0) {
@@ -760,7 +754,7 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
   if ((certlevel>=1 & certlevel!=0) | (certlevel<0.5 & certlevel!=0)){
     stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 80% set certlevel=0.80")
   }
[…]

### 2017-03-22 Cutoff_Algorithm.R
consolidated summarytable and outdata (but not compatible with current cutoff function)

@@ -9,7 +9,7 @@
 
 
 #To complete this process, there are 5 necessary steps
-#I)   Install/load packages and load in functions (run lines 73 - 1350)
+#I)   Install/load packages and load in functions (run lines 67 - 1350)
 #II)  Load in data (run lines: 1370)
 #III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
 #     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
@@ -873,7 +873,7 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
 }
 
 #-----------------------------------------------
-#[IC12] cutuncertgraph
+#[IC12] cutuncertgraph: not yet usable
 #-----------------------------------------------
 #Displays the uncertainty function after components have been combined to create positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
 cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
@@ -932,7 +932,7 @@ cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppressleg
 }
 
 #-----------------------------------------------
-#[IC13]  cutdistgraph
+#[IC13]  cutdistgraph: not yet usable
 #-----------------------------------------------
 #Displays the distributions of the positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively). 
 cutdistgraph<-function(cutobj,pickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
@@ -1041,10 +1041,12 @@ cutdistgraph<-function(cutobj,pickobj,xlim=c(NA,NA),xlab="Optical Density",setbr
 }
 
 #-----------------------------------------------
-#[IC14]  summaryout: combine outputdata and summarytable
+#[IC14]  summaryout: not yet usable
 #-----------------------------------------------
 
-outputdata<-function(standindetobj=NULL,specindetobj=NULL,fileandpathname=NULL){
[…]

### 2017-03-22 Cutoff_Algorithm.R
all graphs corrected, still need to correct summaryout

@@ -30,8 +30,8 @@
 #[IC5]  rawdistgraph
 #[IC6]  rawhistcuts
 #[IC7]  cutoff
-#[IC12] cutuncertgraph
-#[IC13] cutdistgraph
+#[IC8] cutuncertgraph
+#[IC9] cutdistgraph
 #[IC14] summaryout
 #-----------------------------------------------
 #[II] Load and clean data
@@ -861,7 +861,8 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
     cutobj$bound90<-bound90
     cutobj$bound<-bound
   }
-
+  
+  cutobj$par<-modelpickobj$par
   cutobj$uncertobj<-modelpickobj$uncertmat
   cutobj$density<-data.frame(g1=modelpickobj$g1[[cutcomp]], g2=modelpickobj$g2[[cutcomp]])
   cutobj$desc<-modelpickobj$desc
@@ -873,7 +874,7 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
 }
 
 #-----------------------------------------------
-#[IC12] cutuncertgraph: not yet usable
+#[IC8] cutuncertgraph
 #-----------------------------------------------
 #Displays the uncertainty function after components have been combined to create positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
 cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
@@ -916,26 +917,48 @@ cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppressleg
     if (suppresslegend==F){
       legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
       }
-  } else {
-    abline(h =(1-cutobj$desc$uncertlevel), untf = T, col=color[3], lwd=1, lty=2)
+  } else if (cutobj$type=="Non.Standard") {
+    abline(h =(1-cutobj$desc$certlevel), untf = T, col=color[3], lwd=1, lty=2)
[…]

### 2017-03-22 Cutoff_Algorithm.R
Summaryout: adjusting the out portion

@@ -819,6 +819,7 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
     class$group80<-as.factor(class$group80)
     class$group80 = factor(class$group80,levels(class$group80)[c(2,3,1)])
     class$group90 = factor(class$group90,levels(class$group90)[c(2,3,1)])
+    class$group<-NA
   }
   
   if (newcertlevel>0){
@@ -832,6 +833,10 @@ cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
     }
     if (is.na(ubuns)==T){
       message(paste("Upper bound of the ", 100*certlevel, "% indeterminate range not found because uncertainty is never below ", 100*(1-certlevel), "% to the right of the cutpoint. Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.", sep=""))
+    }
+    if (standardcert==F){
+      class$group90<-NA
+      class$group80<-NA
     }
     class$group<-"indeterminate"
     class$group[class$data<=bound[1]]<-"negative"
@@ -1086,27 +1091,26 @@ cutdistgraph<-function(cutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100
 #-----------------------------------------------
 
 #Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 80 and 90 indeterminate ranges and (if desired) one non-standard uncertainty level.
-summaryout<-function(standindetobj=NULL,specindetobj=NULL,fileandpathname=NULL){
+summaryout<-function(cutobj,fileandpathname=NULL){
   
-  if (is.null(standindetobj)==T & is.null(specindetobj)==T) {
-    print("At least one post-cut object must be included")
+  if (is.null(cutobj)==T) {
+    print("Cutoff object must be specified")
+  } 
+  if (cutobj$type=="Standard"){
+    dataout<-cutobj$class[c("id", "data", "group80", "group90")]
+    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]))
+  } else if (cutobj$type=="Non.Standard"){
+    dataout<-cutobj$class[c("id", "data", "group")]
[…]

### 2017-03-22 Cutoff_Algorithm.R
Summary table adjustments

@@ -1097,14 +1097,14 @@ summaryout<-function(cutobj,fileandpathname=NULL){
     print("Cutoff object must be specified")
   } 
   if (cutobj$type=="Standard"){
-    dataout<-cutobj$class[c("id", "data", "group80", "group90")]
-    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]))
+    dataout<-cutobj$class[c("id", "data", "group80", "group90","groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2,5,4,3)]))
   } else if (cutobj$type=="Non.Standard"){
-    dataout<-cutobj$class[c("id", "data", "group")]
-    names(dataout)<-c(names(cutobj$class[c(1,2)]),paste('group', (100*cutobj$desc$certlevel), sep=''))
+    dataout<-cutobj$class[c("id", "data", "group", "groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
   } else {
-    dataout<-cutobj$class[c("id", "data","group80", "group90", "group")]
-    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]),paste('group', (100*cutobj$desc$certlevel), sep=''))
+    dataout<-cutobj$class[c("id", "data","group80", "group90", "group", "groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
   }
 
   if (is.null(fileandpathname)==F){
@@ -1112,8 +1112,10 @@ summaryout<-function(cutobj,fileandpathname=NULL){
   }
 
   outdataobj<-dataout
+
   
   comptable <- function(table){
+    
     dftable<-as.data.frame(table)
     len<-nrow(dftable)
     dftable$orders<-c(NA)[rep(c(1), times=len)]
@@ -1217,11 +1219,11 @@ summaryout<-function(cutobj,fileandpathname=NULL){
     pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
   }
   if (ncol(grps)==4){
-    grpname<-c("Raw Cutoff", paste("Cutoff with ", cutobj$desc$uncertlevel*100, "% Certainty", sep=""))
[…]

### 2017-03-22 PaperEx170308.Rmd
EOD 170321: Functions Fully Consolidated -Need to re-do the table of contents -Need to re-do example 1 -Need to encorporate error handling from Q edits

--- 2017-03-11 PaperEx170308.Rmd
@@
 ```{r loadfunctions, echo=F}
+#Master Cutoff Algorithm Functions
+
+#The cutoff algorithm contained in this R script allows the user to descriminate between a positive and negative sample on a continuous diagnostic without a good gold standard via mixture modeling.
+#The algorithm runs 10 different potential types of mixture models which are based on the normal and skew-normal distributions, looking at 1-5 components each.
+The optimal distribution and component combination is chosen by BIC (which should be minimized)
+#Note: the fitloops function takes some time (typically 2-3 minutes) to run and can fail, if it does, read the error messages and decide whether to try and run it again, or restrict the maximum number of components tested.
+#The user then checks which distribution has been chosen and can manually chose a different distribution if desired (there should be good scientific or clinical reasoning to choose a different combination)
+#If a model with more than two components is optimal by BIC, then the user must allot components to the positive and negative side using their knowlege of the epidemiological conditions where the data comes from.
+#After the decision of where to cut is made, the user can then output graphs and tables that detail what proportion of the population falls into what category, with various degrees of certainty.
+
+#To complete this process, there are 5 necessary steps
+#I)   Install/load packages and load in functions (run lines 67 - 1350)
+#II)  Load in data (run lines: 1370)
+#III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
+#     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
+#     determine which cutpoint is best (if a model with more than two components is optimal) and 3) display the cutoff results
+
+#-----------------------------------------------
+#Table of Contents
+#-----------------------------------------------
+#[I] Install/load packages and load in functions
+#[IA] Install Packages
+#[IB] Load Packages
+#[IC] Load Functions
+#[IC1]  fitloops
+#[IC2]  bicgraph
+#[IC3]  modelpick
+#[IC4]  rawuncertgraph
+#[IC5]  rawdistgraph
+#[IC6]  rawhistcuts
+#[IC7]  cutoff
+
+#-----------------------------------------------
+#[I] Install/load packages and load in functions
+#-----------------------------------------------
+
+#-----------------------------------------------
+#[IA] Install Packages
+#-----------------------------------------------
+
+#install.packages("mixsmsn")
+#install.packages("sn")
 
@@
   loopErr = function(data,i,j,family) {
-    tryCatch(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5)),
+    tryCatch(suppressWarnings(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5))),
              error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
@@
     }
-    fitres$loops<-loops
-    fitres$maxcp<-maxcp
-    fitres$datawithids<-datawithids
-    fitres$errors<-errors
-    return(fitres)
-  } else {
-    if (which.max(errors)<=maxcp){
-      dist<-"Normal"
-    } else {
-      dist<-"Skew.normal"
-    }
-    if (ic<=2){
-      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
-First ensure that the necessary packages have been loaded correctly.
-If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
-    } else {
-      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
-Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
-    }
-  }  
-}
-
-#-----------------------------------------------
-#[IC2]  bestfits
-#-----------------------------------------------
-#Determines the optimal number of distributions and type of distributions (as determined by BIC), was well as the other indicated functions.
-Also creates table of BICs which can be fed into the bicgraph function to visualize which combination of distributions and component numbers is optimal by BIC.
-
-bestfits<-function(fitres){
-
-  maxcp<-fitres$maxcp
-  loops<-fitres$loops
+
     nobj<-vector("list")
@@
     model<-vector("list")
+
+    model$loops<-loops
+    model$maxcp<-maxcp
+    model$datawithids<-datawithids
+    model$errors<-errors
+    model$loopfits<-fitres
+
     model$sn2<-fitres$sn[[2]]
@@
     model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
-  model$datawithids<-fitres$datawithids
-
+
+    print("")
     print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
@@
 
-}
-
-#-----------------------------------------------
-#[IC3]  bicgraph
+  } else {
+    if (which.max(errors)<=maxcp){
+      dist<-"Normal"
+    } else {
+      dist<-"Skew.normal"
+    }
+    if (ic<=2){
+      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
+First ensure that the necessary packages have been loaded correctly.
+If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
+    } else {
+      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
+Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
+    }
+  }
+
+}
+
+#-----------------------------------------------
+#[IC2]  bicgraph
 #-----------------------------------------------
@@
 
-bicgraph<-function(model,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
+bicgraph<-function(fitobj,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
+  model<-fitobj
   color<-setcolor
@@
 #-----------------------------------------------
-#[IC4]  fitpick
+#[IC3]  modelpick
 #-----------------------------------------------
@@
 Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
-
-fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
+The function also performs uncertainty calculations and calculates raw percentages and counts positive and negative for all possible cut points between distributions
+
+modelpick<-function(fitobj, dist="", ncomp=NA){
+  #fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
+  fitres<-fitobj
   if (dist=="" & is.na(ncomp)==T) {
-    dist=bestfits$bestdesc$dist
-    ncomp=bestfits$bestdesc$ncomp
+    dist=fitres$bestdesc$dist
+    ncomp=fitres$bestdesc$ncomp
   }
@@
     stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
-  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
+  } else if ((dist=="Normal" & is.na(fitres$loopfits$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$loopfits$sn[[1]]$bic)==T) | ncomp>length(fitres$loopfits$no) ){
     stop("Must choose a distribution and number of components that were fitted in the fitloops function")
@@
     singlefit<-vector("list")
-    singlefit<-fitres[[dist]][[ncomp]]
+    singlefit<-fitres$loopfits[[dist]][[ncomp]]
 
@@
     for (i in 1:n){
-      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
-      singlefit$sigma2[i]<-fitres[[dist]][[ncomp]]$sigma2[cross$o[i]]
-      singlefit$shape[i]<-fitres[[dist]][[ncomp]]$shape[cross$o[i]]
-      singlefit$pii[i]<-fitres[[dist]][[ncomp]]$pii[cross$o[i]]
-      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
+      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
+      singlefit$sigma2[i]<-fitres$loopfits[[dist]][[ncomp]]$sigma2[cross$o[i]]
+      singlefit$shape[i]<-fitres$loopfits[[dist]][[ncomp]]$shape[cross$o[i]]
+      singlefit$pii[i]<-fitres$loopfits[[dist]][[ncomp]]$pii[cross$o[i]]
+      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
     }
@@
     }
-
-    parameters<-vector("list")
-    parameters$singlefit<-singlefit
-    parameters$par<-par
-    parameters$datawithids<-fitres$datawithids
-    parameters$desc<-vector("list")
-    parameters$desc$ncomp<-ncomp
-    parameters$desc$dist<-distp
-  }
-
-  return(parameters)
-}
-
-#-----------------------------------------------
-#[IC5]  uncert
-#-----------------------------------------------
+  }
+
   #Performs uncertainty calculations.
-
-uncert<-function(fitpickobj){
-
-  data<-fitpickobj$datawithids$data
-  par<-fitpickobj$par
-  fit<-fitpickobj$singlefit
-  n=fitpickobj$desc$ncomp
+  #uncert<-function(fitpickobj){
+
+  fit<-singlefit
   #Create Uncertainties
@@
   for (i in 1:n){
-    if (fitpickobj$desc$dist=="Normal"){
+    if (dist=="Normal"){
       Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
@@
   uncertout<-vector("list")
-  uncertout$uncertmat<-uncertmat
-  uncertout$mcluncert<-mcluncert
-  uncertout$restrmat<-restrmat
-  uncertout$singlefit<-fit
-  uncertout$v<-v
-  uncertout$datawithids<-fitpickobj$datawithids
-  uncertout$desc<-fitpickobj$desc
-  return(uncertout) 
-}
-
-#-----------------------------------------------
-#[IC6]  multcut
-#-----------------------------------------------
-#Calculates raw percentages and counts positive and negative for all possible cut points between distributions
-multcut<-function(uncertobj){
-
-  PD.v<-uncertobj$uncertmat
-  data<-uncertobj$datawithids$data
-  n=uncertobj$desc$ncomp
-  ncomp=n
+
+  #multcut<-function(uncertobj){
+
   closest<-function(uncertdf,min,max,level){
@@
   for (j in 1:(ncomp-1)){
-    uncertdf[[j]]<-data.frame(v=uncertobj$v, uncertainty.v=uncertainty.v[[j]])
+    uncertdf[[j]]<-data.frame(v, uncertainty.v=uncertainty.v[[j]])
     cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
@@
   }
-  class<- data.frame(matrix(unlist(class2), nrow=uncertobj$singlefit$n, byrow=F),stringsAsFactors=TRUE)
+  class<- data.frame(matrix(unlist(class2), nrow=singlefit$n, byrow=F),stringsAsFactors=TRUE)
   class$data<-data
@@
   cutobj<-vector("list")
+
+  cutobj$singlefit<-singlefit
+  cutobj$par<-par
+  cutobj$datawithids<-fitres$datawithids
+  cutobj$desc<-vector("list")
+  cutobj$desc$ncomp<-ncomp
+  cutobj$desc$dist<-distp
+
+  cutobj$uncertmat<-uncertmat
+  cutobj$mcluncert<-mcluncert
+  cutobj$restrmat<-restrmat
+  cutobj$v<-v
+
   cutobj$cutpoint<-cutpoint
@@
   cutobj$cutnames<-grpname
-  #    cutobj$uncertmat<-uncertobj$uncertmat
-  #    cutobj$mcluncert<-uncertobj$mcluncert
-  #    cutobj$restrmat<-uncertobj$restrmat
-  #    cutobj$singlefit<-uncertobj$fit
-  #    cutobj$data<-uncertobj$data
-  #    cutobj$desc<-uncertobj$desc
+
   print(outtab)
@@
 #-----------------------------------------------
-#[IC7]  rawuncertgraph
+#[IC4]  rawuncertgraph
 #-----------------------------------------------
@@
 In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
   "green4",          #1
@@
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(uncertobj$datawithids$data)) 
-    grobj$maxval<-max(uncertobj$datawithids$data)
+    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
+    grobj$maxval<-max(modelpickobj$datawithids$data)
   } else {
@@
 
-  plot(uncertobj$v, uncertobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", uncertobj$desc$dist, "with", uncertobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
+  plot(modelpickobj$v, modelpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
 
   col<-setcolor
-  ncomp<-uncertobj$desc$ncomp
-  for (i in 1:(uncertobj$desc$ncomp-1)){
-    abline(v =multcutobj$cutpoint[i], untf = T, col=col[i], lwd=2)
+  ncomp<-modelpickobj$desc$ncomp
+  for (i in 1:(modelpickobj$desc$ncomp-1)){
+    abline(v =modelpickobj$cutpoint[i], untf = T, col=col[i], lwd=2)
   }
@@
     for (i in 2:ncomp){
-      legtxt[i]<-paste(multcutobj$cutnames[i-1], " , x=", round(multcutobj$cutpoint[i-1], 2), sep="")
+      legtxt[i]<-paste(modelpickobj$cutnames[i-1], " , x=", round(modelpickobj$cutpoint[i-1], 2), sep="")
       colors[i]<-col[i-1]
@@
 #-----------------------------------------------
-#[IC8]  rawdistgraph
+#[IC5]  rawdistgraph
 #-----------------------------------------------
 #Displays the distributions of all the components for the chosen combination of distribution and number of components. 
-rawdistgraph<-function(pickobj,multcutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
+rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
   "green4",          #1
@@
   #Parameter specification
-  ncomp=pickobj$desc$ncomp
+  ncomp=modelpickobj$desc$ncomp
   grobj<-vector("list")
   grobj$gpar<-vector("list")
-  data<-pickobj$datawithids$data
+  data<-modelpickobj$datawithids$data
   color<-setcolor
@@
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data)) 
+    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
     grobj$maxval<-max(data)
@@
   for (i in 1:ncomp){
-    grobj$gpar$mean[i]=pickobj$par$mean[i]
-    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
-    grobj$gpar$shape[i]=pickobj$par$shape[i]
-    grobj$gpar$pii[i]=pickobj$par$pii[i]
+    grobj$gpar$mean[i]=modelpickobj$par$mean[i]
+    grobj$gpar$sd[i]=sqrt(modelpickobj$par$sigma2[i])
+    grobj$gpar$shape[i]=modelpickobj$par$shape[i]
+    grobj$gpar$pii[i]=modelpickobj$par$pii[i]
   }
@@
   for (i in 1:ncomp){
-    if (pickobj$singlefit$shape[i]>0) {
+    if (modelpickobj$singlefit$shape[i]>0) {
       dist[i]<-1
@@
 
-  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, pickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
+  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, modelpickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
   for (i in 1:ncomp) {
@@
 #-----------------------------------------------
-#[IC9]  rawhistcuts
+#[IC6]  rawhistcuts
 #-----------------------------------------------
@@
 
-rawhistcuts<-function(uncertobj,multcutobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
+rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
   "green4",          #1
@@
 )){
-  data<-uncertobj$datawithids$data
+  data<-modelpickobj$datawithids$data
   grobj<-list()
@@
 
-  ncomp<-uncertobj$desc$ncomp
-  cuts <- cut(hist$breaks, c(-Inf,multcutobj$cutpoint,Inf))
+  ncomp<-modelpickobj$desc$ncomp
+  cuts <- cut(hist$breaks, c(-Inf,modelpickobj$cutpoint,Inf))
 
@@
   }
-  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", uncertobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")
+  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", modelpickobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")
 
@@
 #-----------------------------------------------
-#[IC10] standindet
-#-----------------------------------------------
-#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (90 and 95), as well as classification into negative, indeterminate and positive components.
+#[IC7]  cutoff
+#-----------------------------------------------
+#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (80 and 90), as well as classification into negative, indeterminate and positive components.
 It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate.
 In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
-standindet<-function(uncertobj,multcutobj,cutcomp=0){
-  data<-uncertobj$datawithids$data
-  n=uncertobj$desc$ncomp
+cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
+  data<-modelpickobj$datawithids$data
+  n=modelpickobj$desc$ncomp
+  certlevel=newcertlevel
   if (n==2){
@@
   }
-  uncertdf<-multcutobj$uncertdf[[cutcomp]]
+  if ((certlevel>=1 & certlevel!=0) | (certlevel<0.5 & certlevel!=0)){
+    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 80% set certlevel=0.80")
+  }
+  if (standardcert==F & newcertlevel==0){
+    stop("Neither standard nor non-standard certainty levels are requested.
+For standard certainty levels set standardcert=T.
+For non-standard certainty levels set newcertlevel to desired certainty.
+Both standard and nonstandard certainties can be generated in the same function call.")
+  }
+  uncertdf<-modelpickobj$uncertdf[[cutcomp]]
   closest<-function(uncertdf,min,max,level){
@@
   }
-  cutpoint<-multcutobj$cutpoint[cutcomp]  
+
+  cutpoint<-modelpickobj$cutpoint[cutcomp]  
   minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
   minuncub<-closest(uncertdf,cutpoint,max(data),0)
+  class<-data.frame(id=modelpickobj$datawithids$id, data=modelpickobj$datawithids$data)
+  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
+  class$groupdi<-as.factor(class$groupdi)
+
+  if (standardcert==T){
+    lb80<-crosses(uncertdf,minunclb,cutpoint,0.2)  
+    ub80<-crosses(uncertdf,cutpoint,minuncub,0.2)
     lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
     ub90<-crosses(uncertdf,cutpoint,minuncub,0.1)
-  lb95<-crosses(uncertdf,minunclb,cutpoint,0.05)  
-  ub95<-crosses(uncertdf,cutpoint,minuncub,0.05)
-
+    bound80<-c(lb80, ub80)
     bound90<-c(lb90, ub90)
-  bound95<-c(lb95, ub95)
-
+    if (is.na(lb80)==T){
+      message("Lower bound of the 80% indeterminate range not found because uncertainty is never below 20% to the left of the cutpoint.
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
+    }
+    if (is.na(ub80)==T){
+      message("Upper bound of the 80% indeterminate range not found because uncertainty is never below 20% to the right of the cutpoint.
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
+    }
     if (is.na(lb90)==T){
       message("Lower bound of the 90% indeterminate range not found because uncertainty is never below 10% to the left of the cutpoint.
-Consider using non-standard bound function to find more descriptive indeterminate range")
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
     }
@@
       message("Upper bound of the 90% indeterminate range not found because uncertainty is never below 10% to the right of the cutpoint.
-Consider using non-standard bound function to find more descriptive indeterminate range")
-  }
-  if (is.na(lb95)==T){
-    message("Lower bound of the 95% indeterminate range not found because uncertainty is never below 5% to the left of the cutpoint.
-Consider using non-standard bound function to find more descriptive indeterminate range")
-  }
-  if (is.na(ub95)==T){
-    message("Upper bound of the 95% indeterminate range not found because uncertainty is never below 5% to the right of the cutpoint.
-Consider using non-standard bound function to find more descriptive indeterminate range")
-  }
-
-  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
-  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
-  class$group95<-"indeterminate"
-  class$group95[class$data<=bound95[1]]<-"negative"
-  class$group95[class$data>=bound95[2]]<-"positive"
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
+    }
     class$group90<-"indeterminate"
@@
     class$group90[class$data>=bound90[2]]<-"positive"
-
-  class$groupdi<-as.factor(class$groupdi)
-  class$group95<-as.factor(class$group95)
+    class$group80<-"indeterminate"
+    class$group80[class$data<=bound80[1]]<-"negative"
+    class$group80[class$data>=bound80[2]]<-"positive"
     class$group90<-as.factor(class$group90)
-
+    class$group80<-as.factor(class$group80)
+    class$group80 = factor(class$group80,levels(class$group80)[c(2,3,1)])
     class$group90 = factor(class$group90,levels(class$group90)[c(2,3,1)])
-  class$group95 = factor(class$group95,levels(class$group95)[c(2,3,1)])
-
-  cutobj<-vector("list")
-  cutobj$cutpoint<-cutpoint
-  cutobj$bound90<-bound90
-  cutobj$bound95<-bound95
-  cutobj$class<-class
-  cutobj$uncertdf<-uncertdf
-  cutobj$type<-"Standard"
-  cutobj$uncertobj<-uncertobj$uncertmat
-  cutobj$density<-vector("list")
-  cutobj$density$g1<-multcutobj$g1[[cutcomp]]
-  cutobj$density$g2<-multcutobj$g2[[cutcomp]]
-  cutobj$desc<-uncertobj$desc
-  cutobj$desc$cutcomp<-cutcomp
-  cutobj$datawithids<-uncertobj$datawithids
-
-  return(cutobj) 
-}
-
-#-----------------------------------------------
-#[IC11]  specindet
-#-----------------------------------------------
-#Provides cutpoint and allows the user to specify a maximum tolerable uncertianty which will be used to calculate an indeterminate ranges, as well as classification into negative, indeterminate and positive components.
-Maximum tolerable uncertianty is 1-certainty level (ex: uncertlevel=0.05 for 95% indetermintate range since the 95% certainty means that uncertainty of classification is less than 0.05 for all samples classified as positive or negative).
-specindet<-function(uncertobj,multcutobj,certlevel,cutcomp=0){
+    class$group<-NA
+  }
+
+  if (newcertlevel>0){
     uncertlevel<-1-certlevel
     uncertlevelpct<-certlevel
-
-  PD.v<-uncertobj$uncertmat
-  data<-uncertobj$datawithids$data
-  n=uncertobj$desc$ncomp
-
-  if (n==2){
-    cutcomp=1
-  } else if (cutcomp==0) {
-    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
-  }
-  if (certlevel>=1 | certlevel<0.5){
-    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 90% set certlevel=0.90")
-  }
-
-  uncertdf<-multcutobj$uncertdf[[cutcomp]]
-  closest<-function(uncertdf,min,max,level){
-    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
-    diff.cut<-abs(uncertdf2$uncertainty.v - level)
-    min.diffc<-min(diff.cut)
-    y<-which(min.diffc==diff.cut)
-    z<-min(y)
-    cut<-uncertdf2$v[z]  
-    return(cut) 
-  }
-  crosses<-function(uncertdf,min,max,level){
-    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
-    diff.cut<-abs(uncertdf2$uncertainty.v - level)
-    min.diffc<-min(diff.cut)
-    if (min.diffc<=0.001){
-      y<-which(min.diffc==diff.cut)
-      #take first one
-      z<-min(y)
-      cut<-uncertdf2$v[z]  
-    } else {
-      cut<-NA
-    }
-    return(cut)
-  }
-  cutpoint<-closest(uncertdf,min(data),max(data),0.5)
-  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
-  minuncub<-closest(uncertdf,cutpoint,max(data),0)
     lbuns<-crosses(uncertdf,minunclb,cutpoint,uncertlevel)  
@@
     bound<-c(lbuns, ubuns)
-
-  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
-  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
+    if (is.na(lbuns)==T){
+      message(paste("Lower bound of the ", 100*certlevel, "% indeterminate range not found because uncertainty is never below ", 100*(1-certlevel), "% to the left of the cutpoint.
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.", sep=""))
+    }
+    if (is.na(ubuns)==T){
+      message(paste("Upper bound of the ", 100*certlevel, "% indeterminate range not found because uncertainty is never below ", 100*(1-certlevel), "% to the right of the cutpoint.
+Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.", sep=""))
+    }
+    if (standardcert==F){
+      class$group90<-NA
+      class$group80<-NA
+    }
     class$group<-"indeterminate"
@@
     class$group[class$data>=bound[2]]<-"positive"
-
-  class$groupdi<-as.factor(class$groupdi)
     class$group<-as.factor(class$group)
     class$group = factor(class$group,levels(class$group)[c(2,3,1)])
+  }
 
@@
   cutobj$cutpoint<-cutpoint
-  cutobj$bound<-bound
+  cutobj$bound80<-c(NA,NA)
+  cutobj$bound90<-c(NA,NA)
+  cutobj$bound<-c(NA,NA)
   cutobj$class<-class
   cutobj$uncertdf<-uncertdf
+
+  if (standardcert==T & newcertlevel==0){
+    cutobj$type<-"Standard"
+    cutobj$bound80<-bound80
+    cutobj$bound90<-bound90
+  } else if (standardcert==F & newcertlevel>0) {
     cutobj$type<-"Non.Standard"
-  cutobj$uncertobj<-uncertobj$uncertmat
-  cutobj$density<-data.frame(g1=multcutobj$g1[[cutcomp]], g2=multcutobj$g2[[cutcomp]])
-  cutobj$desc<-uncertobj$desc
+    cutobj$bound<-bound
+  } else {
+    cutobj$type<-"Standard.and.Non"
+    cutobj$bound80<-bound80
+    cutobj$bound90<-bound90
+    cutobj$bound<-bound
+  }
+
+  cutobj$par<-modelpickobj$par
+  cutobj$uncertobj<-modelpickobj$uncertmat
+  cutobj$density<-data.frame(g1=modelpickobj$g1[[cutcomp]], g2=modelpickobj$g2[[cutcomp]])
+  cutobj$desc<-modelpickobj$desc
   cutobj$desc$cutcomp<-cutcomp
-  cutobj$desc$uncertlevel<-uncertlevelpct
-  cutobj$datawithids<-uncertobj$datawithids
+  cutobj$desc$certlevel<-newcertlevel
+  cutobj$datawithids<-modelpickobj$datawithids
 
@@
 #-----------------------------------------------
-#[IC12]  cutuncertgraph
+#[IC8] cutuncertgraph
 #-----------------------------------------------
@@
   color=setcolor
+  grobj<-vector("list")
 
@@
   if (cutobj$type=="Standard") {
-    abline(h =0.05, untf = T, col=color[4], lwd=1, lty=2)
-    abline(h =0.1, untf = T, col=color[2], lwd=1, lty=2)
-    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
+    abline(h =0.1, untf = T, col=color[4], lwd=1, lty=2)
+    abline(h =0.2, untf = T, col=color[2], lwd=1, lty=2)
+    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
+    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
     legtxt<-c("Uncertainty",
               paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
+              "Uncertainty=0.20",
+              "Indeterminate Range with", 
+              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
+              "Uncertainty=0.10",
+              "Indeterminate Range with", 
+              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep ="")
+    )
+    if (suppresslegend==F){
+      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
+      }
+  } else if (cutobj$type=="Non.Standard") {
+    abline(h =(1-cutobj$desc$certlevel), untf = T, col=color[3], lwd=1, lty=2)
+    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
+    legtxt<-c("Uncertainty",
+              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
+              paste("Uncertainty=", (1-cutobj$desc$certlevel), sep =""),
+              "Indeterminate Range with", 
+              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
+    )
+    if (suppresslegend==F){
+      legend("topright", col=c(color[9],color[6], color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1), legend=legtxt, xjust=1, seg.len=1, title="Cutpoint, Uncertainties, and Boundaries")
+    }
+  } else if (cutobj$type=="Standard.and.Non") {
+    abline(h =0.1, untf = T, col=color[4], lwd=1, lty=2)
+    abline(h =0.2, untf = T, col=color[2], lwd=1, lty=2)
+    abline(h =(1-cutobj$desc$certlevel), untf = T, col=color[3], lwd=1, lty=2)
+    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
+    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
+    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
+    legtxt<-c("Uncertainty",
+              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
+              "Uncertainty=0.20",
+              "Indeterminate Range with", 
+              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
               "Uncertainty=0.10",
@@
               paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
-              "Uncertainty=0.05",
+              paste("Uncertainty=", (1-cutobj$desc$uncertlevel), sep =""),
               "Indeterminate Range with", 
-              paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
+              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
     )
     if (suppresslegend==F){
-      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
-      }
-  } else {
-    abline(h =(1-cutobj$desc$uncertlevel), untf = T, col=color[3], lwd=1, lty=2)
-    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
-    legtxt<-c("Uncertainty",
-              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-              paste("Uncertainty=", (1-cutobj$desc$uncertlevel), sep =""),
-              "Indeterminate Range with", 
-              paste((100*cutobj$desc$uncertlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
-    )
-    if (suppresslegend==F){
-      legend("topright", col=c(color[9],color[6], color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1), legend=legtxt, xjust=1, seg.len=1, title="Cutpoint, Uncertainties, and Boundaries")
-    }
-  }
-}
-
-#-----------------------------------------------
-#[IC13]  cutdistgraph
+      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white", color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
+    }
+  }
+}
+
+#-----------------------------------------------
+#[IC9]  cutdistgraph
 #-----------------------------------------------
@@
 Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
-cutdistgraph<-function(cutobj,pickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
+cutdistgraph<-function(cutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
   "green4",          #1
@@
   cutobj<-cutobj
-  data<-pickobj$datawithids$data
+  data<-cutobj$datawithids$data
   color<-setcolor
@@
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) { 
-    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data))     
+    grobj$minval<-(1/10)*trunc(10*min(data))     
     grobj$maxval<-max(data)
@@
   for (i in 1:ncomp){
-    grobj$gpar$mean[i]=pickobj$par$mean[i]
-    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
-    grobj$gpar$shape[i]=pickobj$par$shape[i]
-    grobj$gpar$pii[i]=pickobj$par$pii[i]
+    grobj$gpar$mean[i]=cutobj$par$mean[i]
+    grobj$gpar$sd[i]=sqrt(cutobj$par$sigma2[i])
+    grobj$gpar$shape[i]=cutobj$par$shape[i]
+    grobj$gpar$pii[i]=cutobj$par$pii[i]
   }
@@
   grobj$ylim<-c(0, max(grobj$gt[[2+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
-  grobj$xlim<-c(round(min(pickobj$datawithids$data),digits=1),grobj$maxval)
+  grobj$xlim<-c(round(min(data),digits=1),grobj$maxval)
 
@@
   if (cutobj$type=="Standard") {
-    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
+    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
+    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
     if (suppresslegend==F) {
@@
                 "Indeterminate Range with", 
-                paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
+                paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
                 "Indeterminate Range with", 
-                paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
+                paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep ="")
       )
@@
     }
-  } else {
+  } else if (cutobj$type=="Non.Standard") {
     abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
@@
     }
+  } else {
+    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
+    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
+    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
+    if (suppresslegend==F){
+    legtxt<-c("Negative Distribution",
+              "Positive Distribution",
+              "Overall Distribution",
+              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
+              "Indeterminate Range with", 
+              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
+              "Indeterminate Range with", 
+              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
+              "Indeterminate Range with", 
+              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
+      )
+    legend("topright", col=c(color[1], color[5], color[9], color[6], color[2], "white", color[4], "white", color[3], "white"), lwd=c(2,2,2,1,1,1,1,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1,1,1,1,1))
+    }
   }
@@
 #-----------------------------------------------
-#[IC14]  outputdata
-#-----------------------------------------------
-outputdata<-function(standindetobj=NULL,specindetobj=NULL,fileandpathname=NULL){
-  if (is.null(standindetobj)==T & is.null(specindetobj)==T) {
-    print("At least one post-cut object must be included")
-  } else {
-    if (is.null(standindetobj)==T) {
-      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
-      dataout<-specindetobj$class
-    } else if (is.null(specindetobj)==T){
-      dataout<-standindetobj$class
-    } else {
-      datastand<-standindetobj$class[c("id", "group90", "group95")]
-      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
-      datanonst<-specindetobj$class
-      dataout<-merge(datanonst,datastand,by.x="id", by.y="id")
-    }
+#[IC10]  summaryout
+#-----------------------------------------------
+
+#Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 80 and 90 indeterminate ranges and (if desired) one non-standard uncertainty level.
+summaryout<-function(cutobj,fileandpathname=NULL){
+
+  if (is.null(cutobj)==T) {
+    print("Cutoff object must be specified")
+  } 
+  if (cutobj$type=="Standard"){
+    dataout<-cutobj$class[c("id", "data", "group80", "group90","groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2,5,4,3)]))
+  } else if (cutobj$type=="Non.Standard"){
+    dataout<-cutobj$class[c("id", "data", "group", "groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
+  } else {
+    dataout<-cutobj$class[c("id", "data","group80", "group90", "group", "groupdi")]
+    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
+  }
+
   if (is.null(fileandpathname)==F){
@@
   }
-  }
-  return(dataout)
-}
-
-#-----------------------------------------------
-#[IC15]  summarytable
-#-----------------------------------------------
-#Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 90 and 95 indeterminate ranges and (if desired) one non-standard uncertainty level.
-summarytable<-function(outdataobj,standindetobj=NULL,specindetobj=NULL){
+
+  outdataobj<-dataout
+
   comptable <- function(table){
+
     dftable<-as.data.frame(table)
@@
   if (ncol(outdataobj)!=4 ){ 
+    grps80<-table(outdataobj$group80)
+    grps80<-comptable(grps80)
+    grps$group80<-grps80$Freq
     grps90<-table(outdataobj$group90)
@@
     grps$group90<-grps90$Freq
-    grps95<-table(outdataobj$group95)
-    grps95<-comptable(grps95)
-    grps$group95<-grps95$Freq
   }
@@
 
-  vars<-c("groupdi", "group90", "group95", "group80")
+  vars<-c("groupdi", "group80", "group90", "group80")
 
@@
   if (ncol(grps)==4){
-    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""))
+    grpname<-c("Raw Cutoff", paste("Cutoff with ", cutobj$desc$certlevel*100, "% Certainty", sep=""))
   } else if (ncol(grps)==5) {
-    grpname<-c("Raw Cutoff", "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
+    grpname<-c("Raw Cutoff", "Cutoff with 80% Certainty", "Cutoff with 90% Certainty")
   } else if (ncol(grps)==6) {
-    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""), "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
+    grpname<-c("Raw Cutoff", paste("Cutoff with ", cutobj$desc$certlevel*100, "% Certainty", sep=""), "Cutoff with 80% Certainty", "Cutoff with 90% Certainty")
   }
@@
   model<-vector("list")
-  if (is.null(specindetobj)==F){
-    model$cutpoint<-specindetobj$cutpoint
-    model$desc<-specindetobj$desc
+  if (is.null(cutobj)==F){
+    model$cutpoint<-cutobj$cutpoint
+    model$desc<-cutobj$desc
     model$bounds<-vector("list")
-    model$bounds$boundns<-specindetobj$bound
-  } else {
-    model$cutpoint<-standindetobj$cutpoint     
-    model$desc<-standindetobj$desc
+    model$bounds$boundns<-cutobj$bound
+  } else {
+    model$cutpoint<-cutobj$cutpoint     
+    model$desc<-cutobj$desc
     model$bounds<-vector("list")
-    model$bounds$bound90<-standindetobj$bound90
-    model$bounds$bound95<-standindetobj$bound95
-  }
-  if (is.null(standindetobj)==F & is.null(specindetobj)==F){
-    model$bounds$bound90<-standindetobj$bound90
-    model$bounds$bound95<-standindetobj$bound95
+    model$bounds$bound80<-cutobj$bound80
+    model$bounds$bound90<-cutobj$bound90
+  }
+  if (is.null(cutobj)==F & is.null(cutobj)==F){
+    model$bounds$bound80<-cutobj$bound80
+    model$bounds$bound90<-cutobj$bound90
   }
@@
 
-  message(paste("The absolute cutoff is", round(standindetobj$cutpoint,3), sep=" "))
+  message(paste("The absolute cutoff is", round(cutobj$cutpoint,3), sep=" "))
   message("Boundries of Indeterminates:")
-  if (is.null(specindetobj)==F) {
-    message(paste(specindetobj$desc$uncertlevel*100, "% Certainty Range: (", round(specindetobj$bound[1], 3), ", ", round(specindetobj$bound[2], 3), ")", sep=""))
-  }
-  message(paste("90% Certainty Range: (", round(standindetobj$bound90[1], 3), ", ", round(standindetobj$bound90[2], 3), ")", sep=""))
-  message(paste("95% Certainty Range: (", round(standindetobj$bound95[1], 3), ", ", round(standindetobj$bound95[2], 3), ")", sep=""))
+  if (cutobj$type!="Standard"){
+    message(paste(cutobj$desc$certlevel*100, "% Certainty Range: (", round(cutobj$bound[1], 3), ", ", round(cutobj$bound[2], 3), ")", sep=""))
+  }
+  if (cutobj$type!="Non.Standard"){  
+    message(paste("80% Certainty Range: (", round(cutobj$bound80[1], 3), ", ", round(cutobj$bound80[2], 3), ")", sep=""))
+    message(paste("90% Certainty Range: (", round(cutobj$bound90[1], 3), ", ", round(cutobj$bound90[2], 3), ")", sep=""))
+  }
 
@@
 
-#-----------------------------------------------
-#[IC16]  doallprobsfail
-#-----------------------------------------------
-#Performs all of the above functions, but is very unlikely to function since in the majority of cases a the two component case will not be optimal by BIC
-doallprobsfail<-function(datawithids,fitres=NA){
-
-  if (length(fitres)==1){
-    fitres5<-fitloops(datawithids)
-  }
-
-  bestfit5<-bestfits(fitres5)
-  bicgraph(bestfit5)
-  pickobj5<-fitpick(fitres5,bestfit5)
-  uncertobj5<-uncert(pickobj5)
-  multcutobj5<-multcut(uncertobj5)
-  rawdistgraph(pickobj5)
-  rawuncertgraph(uncertobj5,multcutobj5)
-  rawhistcuts(uncertobj5,multcutobj5)
-  standindetpts5<-standindet(uncertobj5,multcutobj5)
-  cutuncertgraph(standindetpts5)
-  cutdistgraph(standindetpts5, pickobj5)
-  nsindet5<-nsstandindets(uncertobj5,multcutobj5,certlevel=0.8,cutcomp=1)
-  cutuncertgraph(nsindet5)
-  cutdistgraph(nsindet5, pickobj5)
-  outdata5<-outputdata(standindetpts5,nsindet5)
-  summary<-summarytable(outdata5, standindetpts5, nsindet5)
-
-}
-
-#-----------------------------------------------
-#[IC17]  twobestdorest
-#-----------------------------------------------
-#Performs all functions after bestfit if the optimal number of components is 2.
-twobestdorest<-function(fitloopsobj,bestfit){
-
-  pickobj<-fitpick(fitloopsobj,bestfit)
-  uncertobj<-uncert(pickobj)
-  multcutobj<-multcut(uncertobj)
-  rawdistgraph(pickobj)
-  rawuncertgraph(uncertobj,multcutobj)
-  rawhistcuts(uncertobj,multcutobj)
-  scptsobj<-standindet(uncertobj,multcutobj)
-  cutuncertgraph(scptsobj)
-  cutdistgraph(scptsobj, pickobj)
-  outdata<-outputdata(standindetpts,nsindet)
-  summary<-summarytable(outdata, standindetpts, nsindet)
-
-  model<-vector("list")
-  model$uncertobj<-uncertobj
-  model$multcutobj<-multcutobj
-  model$scptsobj<-scptsobj
-  model$outdataobj<-outdata
-  model$summaryobj<-summary
-
-  return(model)
-
-}
 ```
@@
 
-examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))
-
-hist(examp1$data, breaks=100, main="Histogram of Data")
-dens1<-density(examp1$data)
-plot(dens1, main="Density of Data")
+#examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))
+
+#hist(examp1$data, breaks=100, main="Histogram of Data")
+#dens1<-density(examp1$data)
+#plot(dens1, main="Density of Data")
 
@@
   #might not run due to 5 component solution not being easily optimized, if it doesn't run either load the full file, or run fitloopsobject<-fitloops(datawithids=examp2,maxcp=4): 
-load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")
-
-fitloopsobject1<-ex1
-
-bestfitsobject1<-bestfits(fitres=fitloopsobject1)
-bicgraph(model=bestfitsobject1)
-
-fitpickobject1<-fitpick(fitres=fitloopsobject1,bestfits=bestfitsobject1)
-
-uncertobject1<-uncert(fitpickobj=fitpickobject1)
-multcutobject1<-multcut(uncertobj=uncertobject1)
-
-rawuncertgraph(uncertobj=uncertobject1,multcutobj=multcutobject1)
-rawdistgraph(pickobj=fitpickobject1)
-rawhistcuts(uncertobj=uncertobject1,multcutobj=multcutobject1)
-
-standindetobject1<-standindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1)
-specindetobject1<-specindet(uncertobj=uncertobject1,multcutobj=multcutobject1,cutcomp=1,certlevel=0.8)
-
-cutuncertgraph(cutobj=standindetobject1)
-cutdistgraph(cutobj=standindetobject1, pickobj=fitpickobject1)
-cutuncertgraph(cutobj=specindetobject1)
-cutdistgraph(cutobj=specindetobject1, pickobj=fitpickobject1)
-outdataobject1<-outputdata(standindet=standindetobject1,specindet=specindetobject1)
-summaryobject1<-summarytable(outdataobj=outdataobject1, standindetobj=standindetobject1, specindetobj=specindetobject1)
-outtable1<-summaryobject1$outtab
-outtable1
+#load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")
+
+#bicgraph(model=fitloopsobject1)
+
+#modelpickobject2<-modelpick(fitloopsobject1)
+
+#rawuncertgraph(modelpickobject1)
+#rawdistgraph(modelpickobject1)
+#rawhistcuts(modelpickobject1)
 
@@
 
-fitloopsobject2<-fitloops(datawithids=examp2)
-bestfitsobject2<-bestfits(fitres=fitloopsobject2)
-bicgraph(model=bestfitsobject2)
-
-fitpickobject2<-fitpick(fitres=fitloopsobject2,bestfits=bestfitsobject2)
-
-uncertobject2<-uncert(fitpickobj=fitpickobject2)
-multcutobject2<-multcut(uncertobj=uncertobject2)
-
-rawuncertgraph(uncertobj=uncertobject2,multcutobj=multcutobject2)
-rawdistgraph(pickobj=fitpickobject2)
-rawhistcuts(uncertobj=uncertobject2,multcutobj=multcutobject2)
-
-standindetobject2<-standindet(uncertobj=uncertobject2,multcutobj=multcutobject2,cutcomp=1)
-specindetobject2<-specindet(uncertobj=uncertobject2,multcutobj=multcutobject2,cutcomp=1,certlevel=0.8)
-
-cutuncertgraph(cutobj=standindetobject2, xlim=c(min(examp2$data),3))
-cutdistgraph(cutobj=standindetobject2, pickobj=fitpickobject2, xlim=c(min(examp2$data),3))
-cutuncertgraph(cutobj=specindetobject2, xlim=c(min(examp2$data),3))
-cutdistgraph(cutobj=specindetobject2, pickobj=fitpickobject2, xlim=c(min(examp2$data),3))
-outdataobject2<-outputdata(standindet=standindetobject2,specindet=specindetobject2)
-summaryobject2<-summarytable(outdataobj=outdataobject2, standindetobj=standindetobject2, specindetobj=specindetobject2)
-outtable2<-summaryobject2$outtab
-outtable2
-
+fitloopsobject2<-fitloops(datawithids = examp2)
+
+bicgraph(fitobj = fitloopsobject2)
+
+modelpickobject2<-modelpick(fitobj = fitloopsobject2)
+
+rawuncertgraph(modelpickobj = modelpickobject2)
+rawdistgraph(modelpickobj = modelpickobject2)
+rawhistcuts(modelpickobj = modelpickobject2)
+
+cutoffobject2<-cutoff(modelpickobj = modelpickobject2, cutcomp = 1)
+
+cutuncertgraph(cutobj = cutoffobject2)
+cutdistgraph(cutobj = cutoffobject2)
+
+summaryobject2<-summaryout(cutobj = cutoffobject2)
 ```

### 2017-03-22 Cutoff_Algorithm.R

@@ -9,9 +9,9 @@
 
 
 #To complete this process, there are 5 necessary steps
-#I)   Install/load packages and load in functions (run lines 67 - 1350)
-#II)  Load in data (run lines: 1370)
-#III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
+#I)   Install/load packages and load in functions (run lines 67 - 1270)
+#II)  Load in data (run lines: 1288)
+#III) Run each line of code separately, from 1314 to the end of the program, following the directions in the comments
 #     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
 #     determine which cutpoint is best (if a model with more than two components is optimal) and 3) display the cutoff results
  
@@ -32,7 +32,7 @@
 #[IC7]  cutoff
 #[IC8] cutuncertgraph
 #[IC9] cutdistgraph
-#[IC14] summaryout
+#[IC10] summaryout
 #-----------------------------------------------
 #[II] Load and clean data
 #[IIA] Import data 
@@ -283,7 +283,8 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
 #-----------------------------------------------
 #Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.
 
-bicgraph<-function(model,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
+bicgraph<-function(fitobj,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
+  model<-fitobj
   color<-setcolor
   buffer<-(max(model$bictab)-min(model$bictab))/20
   ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
@@ -305,8 +306,9 @@ bicgraph<-function(model,title="BIC by type and number of components",setcolor=c
 #-----------------------------------------------
[…]

### 2017-03-22 PaperEx170308.Rmd
Cleaned data (removed duplicates), updated cutoff algorithm, and updated analyses for paper

@@
   dists="Both"
-  datawithids<-unique(datawithids)
+  datawithids2<-unique(datawithids)
+  if (nrow(unique(datawithids))!=nrow(datawithids)){
+    stop("The input data (datawithids) must not include duplicates. Please remove duplicates.")
+  }
   data<-datawithids$data
@@
 
-#examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))
-
-#hist(examp1$data, breaks=100, main="Histogram of Data")
-#dens1<-density(examp1$data)
-#plot(dens1, main="Density of Data")
-
-#fitloopsobject<-fitloops(datawithids=examp2)
-  #might not run due to 5 component solution not being easily optimized, if it doesn't run either load the full file, or run fitloopsobject<-fitloops(datawithids=examp2,maxcp=4): 
-#load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")
-
-#bicgraph(model=fitloopsobject1)
-
-#modelpickobject2<-modelpick(fitloopsobject1)
-
-#rawuncertgraph(modelpickobject1)
-#rawdistgraph(modelpickobject1)
-#rawhistcuts(modelpickobject1)
+examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))
+
+hist(examp1$data, breaks=100, main="Histogram of Data")
+dens1<-density(examp1$data)
+plot(dens1, main="Density of Data")
+
+#fitloopsobject1<-fitloops(datawithids=examp2)
+  #might not run due to 5 component solution not being easily optimized, if it doesn't run either load the full file, or run fitloopsobject1<-fitloops(datawithids=examp2,maxcp=4): 
+load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")
+
+bicgraph(fitobj=fitloopsobject1)
+
+modelpickobject1<-modelpick(fitloopsobject1)
+
+rawuncertgraph(modelpickobject1)
+rawdistgraph(modelpickobject1)
+rawhistcuts(modelpickobject1)
+
+cutoffobject1<-cutoff(modelpickobj = modelpickobject1)
+
+cutuncertgraph(cutobj = cutoffobject1)
+cutdistgraph(cutobj = cutoffobject1)
+
+summaryobject1<-summaryout(cutobj = cutoffobject1)
+summaryobject1$outtab
 
@@
 
-examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv")
+examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", stringsAsFactors = F, colClasses=c("NULL", NA, NA))
+examp2<-examp2[,c("id", "data")]
 
@@
 summaryobject2<-summaryout(cutobj = cutoffobject2)
+summaryobject2$outtab
 ```

### 2017-03-22 Cutoff_Algorithm.R

@@ -87,7 +87,10 @@ library("sn")
 #accepts a dataframe with columns id and data
 fitloops<-function(datawithids,loops=10,maxcp=5){
   dists="Both"
-  datawithids<-unique(datawithids)
+  datawithids2<-unique(datawithids)
+  if (nrow(unique(datawithids))!=nrow(datawithids)){
+    stop("The input data (datawithids) must not include duplicates. Please remove duplicates.")
+  }
   data<-datawithids$data
   datapts<-length(data)
   norms<-vector("list")
@@ -139,7 +142,7 @@ fitloops<-function(datawithids,loops=10,maxcp=5){
       ic=i
     }
   }
-  
+
   if (i==maxcp & j==loops+1 & max(errors[maxcp, 2])<=2 & max(errors[maxcp, 1])<=2 ) {
     nobj<-vector("list")
     nobj$mc<-matrix(nrow=maxcp, ncol=loops)
@@ -1326,7 +1329,7 @@ plot(dens, main="Density of Data")
 
 fitloopsobject<-fitloops(datawithids = examp)
 
-bicgraph(fitobj = fitloopsobject2)
+bicgraph(fitobj = fitloopsobject)
 
 #-----------------------------------------------
 #[IIIC] Pick the best combination of distribution and components (usually the one that is optimal by BIC unless there is scientific rationale)
@@ -1340,7 +1343,7 @@ bicgraph(fitobj = fitloopsobject2)
 #                 ncomp: number from 1 to 5
 #-----------------------------------------------
 
-modelpickobject2<-modelpick(fitobj = fitloopsobject2)
+modelpickobject<-modelpick(fitobj = fitloopsobject)
 
 
 #-----------------------------------------------
@@ -1352,9 +1355,9 @@ modelpickobject2<-modelpick(fitobj = fitloopsobject2)
 #-----------------------------------------------
 #[IVB] Utilize visualizations of uncertainty, distributions and cutpoint to aid in decision-making
 #-----------------------------------------------
-rawuncertgraph(modelpickobj = modelpickobject2)
-rawdistgraph(modelpickobj = modelpickobject2)
-rawhistcuts(modelpickobj = modelpickobject2)
+rawuncertgraph(modelpickobj = modelpickobject)
+rawdistgraph(modelpickobj = modelpickobject)
+rawhistcuts(modelpickobj = modelpickobject)
 
 #-----------------------------------------------
[…]

### 2017-03-22 PaperEx170308.Rmd
Add option to change title

@@
 In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
@@
 
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " ")
+  }
   grobj<-list()
@@
 
-  plot(modelpickobj$v, modelpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
+  plot(modelpickobj$v, modelpickobj$mcluncert, main=title, type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
 
@@
 #Displays the distributions of all the components for the chosen combination of distribution and number of components. 
-rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
+rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
@@
 
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste("Distribution of", ncomp, modelpickobj$desc$dist,"Components", sep = " ")
+  }
+
   grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)
@@
 
-  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, modelpickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
+  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=title, border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
   for (i in 1:ncomp) {
@@
 
-rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
+rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
@@
   ncomp<-modelpickobj$desc$ncomp
+
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste("Classification by", modelpickobj$desc$dist, "with", ncomp, "components", sep = " ")
+  }
+
   cuts <- cut(hist$breaks, c(-Inf,modelpickobj$cutpoint,Inf))
@@
   }
-  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", modelpickobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")
+  plot(hist, col=colors, xlab=xlab, ylab="Density", main=title, freq=F, lty="blank")
 
@@
 Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
-cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+cutuncertgraph<-function(cutobj, xlab="Optical Density", xlim=c(NA,NA), suppresslegend=F, title="default", setcolor=c(
   "green4",          #1
@@
 
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " ")
+  }
+
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty")
+    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=title, type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty")
   } else {
@@
     grobj$maxval<-xlim[2]
-    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty", xlim=c(grobj$minval,grobj$maxval))
+    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=title, type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty", xlim=c(grobj$minval,grobj$maxval))
   }
@@
 Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
-cutdistgraph<-function(cutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
+cutdistgraph<-function(cutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
@@
 
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste(cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " ")
+  }
+
   grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)
@@
   #Plot Graph
-  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste(cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "), border=color[8], ylim=grobj$ylim, xlim=grobj$xlim)
+  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=title, border=color[8], ylim=grobj$ylim, xlim=grobj$xlim)
 
@@
                 "Indeterminate Range with", 
-                paste((cutobj$desc$uncertlevel*100), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
+                paste((cutobj$desc$certlevel*100), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
       )
@@
 
-hist(examp1$data, breaks=100, main="Histogram of Data")
+hist(examp1$data, breaks=100, main="")
 dens1<-density(examp1$data)
-plot(dens1, main="Density of Data")
+plot(dens1, main="")
 
@@
 
-bicgraph(fitobj=fitloopsobject1)
+bicgraph(fitobj=fitloopsobject1, title="")
 
@@
 
-rawuncertgraph(modelpickobject1)
-rawdistgraph(modelpickobject1)
-rawhistcuts(modelpickobject1)
+rawuncertgraph(modelpickobject1, title="")
+rawdistgraph(modelpickobject1, title="")
+rawhistcuts(modelpickobject1, title="")
 
@@
 
-cutuncertgraph(cutobj = cutoffobject1)
-cutdistgraph(cutobj = cutoffobject1)
+cutuncertgraph(cutobj = cutoffobject1, title="")
+cutdistgraph(cutobj = cutoffobject1, title="")
 
@@
 
-hist(examp2$data, breaks=100, main="Histogram of Data")
+hist(examp2$data, breaks=100, main="")
 dens2<-density(examp2$data)
-plot(dens2, main="Density of Data")
+plot(dens2, main="")
 
@@
 
-bicgraph(fitobj = fitloopsobject2)
+bicgraph(fitobj = fitloopsobject2, title="")
 
@@
 
-rawuncertgraph(modelpickobj = modelpickobject2)
-rawdistgraph(modelpickobj = modelpickobject2)
-rawhistcuts(modelpickobj = modelpickobject2)
+rawuncertgraph(modelpickobj = modelpickobject2, title="")
+rawdistgraph(modelpickobj = modelpickobject2, title="")
+rawhistcuts(modelpickobj = modelpickobject2, title="")
 
@@
 
-cutuncertgraph(cutobj = cutoffobject2)
-cutdistgraph(cutobj = cutoffobject2)
+cutuncertgraph(cutobj = cutoffobject2, title="")
+cutdistgraph(cutobj = cutoffobject2, title="")
 

### 2017-03-22 Cutoff_Algorithm.R

@@ -543,7 +543,7 @@ modelpick<-function(fitobj, dist="", ncomp=NA){
 #[IC4]  rawuncertgraph
 #-----------------------------------------------
 #Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
+rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
   "turquoise3",      #2
   "royalblue2",      #3
@@ -555,6 +555,11 @@ rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegen
   "black"            #9
 )){
   
+  if (title!="default"){
+    title=title
+  } else {
+    title=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " ")
+  }
   grobj<-list()
   if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
     grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
@@ -564,7 +569,7 @@ rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegen
     grobj$maxval<-xlim[2]
   }
 
-  plot(modelpickobj$v, modelpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
+  plot(modelpickobj$v, modelpickobj$mcluncert, main=title, type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
[…]

### 2017-03-22 PaperEx170308.Rmd
Cutoff_Algorithm file now only contains function definitions, and the example remains but has been commented out. As such it can be called in from other documents to solely use the functions

--- 2017-03-22 PaperEx170308.Rmd
@@
 #Laptop
-opts_knit$set(root.dir = "/Users/sarahsullivan/Google Drive/Sarah Sullivan/Code")
+opts_knit$set(root.dir = "/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/")
 ```
@@
 ```{r loadfunctions, echo=F}
-#Master Cutoff Algorithm Functions
-
-#The cutoff algorithm contained in this R script allows the user to descriminate between a positive and negative sample on a continuous diagnostic without a good gold standard via mixture modeling.
-#The algorithm runs 10 different potential types of mixture models which are based on the normal and skew-normal distributions, looking at 1-5 components each.
-The optimal distribution and component combination is chosen by BIC (which should be minimized)
-#Note: the fitloops function takes some time (typically 2-3 minutes) to run and can fail, if it does, read the error messages and decide whether to try and run it again, or restrict the maximum number of components tested.
-#The user then checks which distribution has been chosen and can manually chose a different distribution if desired (there should be good scientific or clinical reasoning to choose a different combination)
-#If a model with more than two components is optimal by BIC, then the user must allot components to the positive and negative side using their knowlege of the epidemiological conditions where the data comes from.
-#After the decision of where to cut is made, the user can then output graphs and tables that detail what proportion of the population falls into what category, with various degrees of certainty.
-
-#To complete this process, there are 5 necessary steps
-#I)   Install/load packages and load in functions (run lines 67 - 1350)
-#II)  Load in data (run lines: 1370)
-#III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
-#     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
-#     determine which cutpoint is best (if a model with more than two components is optimal) and 3) display the cutoff results
-
-#-----------------------------------------------
-#Table of Contents
-#-----------------------------------------------
-#[I] Install/load packages and load in functions
-#[IA] Install Packages
-#[IB] Load Packages
-#[IC] Load Functions
-#[IC1]  fitloops
-#[IC2]  bicgraph
-#[IC3]  modelpick
-#[IC4]  rawuncertgraph
-#[IC5]  rawdistgraph
-#[IC6]  rawhistcuts
-#[IC7]  cutoff
-
-#-----------------------------------------------
-#[I] Install/load packages and load in functions
-#-----------------------------------------------
-
-#-----------------------------------------------
-#[IA] Install Packages
-#-----------------------------------------------
-
-#install.packages("mixsmsn")
-#install.packages("sn")
-
-#-----------------------------------------------
-#[IB] Load Packages
-#-----------------------------------------------
-
-library("mixsmsn")
-library("sn")
-
-#-----------------------------------------------
-#[IC] Load Functions
-#-----------------------------------------------
-
-#-----------------------------------------------
-#[IC1]  fitloops
-#-----------------------------------------------
-
-#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution.
-This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively).
-This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails.
-The default values are to do both skew-normal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components.
-It is not reccommended to change these unless the default function fails.
-#accepts a dataframe with columns id and data
-fitloops<-function(datawithids,loops=10,maxcp=5){
-  dists="Both"
-  datawithids2<-unique(datawithids)
-  if (nrow(unique(datawithids))!=nrow(datawithids)){
-    stop("The input data (datawithids) must not include duplicates. Please remove duplicates.")
-  }
-  data<-datawithids$data
-  datapts<-length(data)
-  norms<-vector("list")
-  errors<-matrix(0, nrow=maxcp, ncol=2)
-  loopErr = function(data,i,j,family) {
-    tryCatch(suppressWarnings(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5))),
-             error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
-               NA}) 
-  }
-
-  for (i in 1:maxcp){
-    norms$mc[[i]]<-vector("list", maxcp)
-    norms$ms[[i]]<-vector("list", maxcp)
-    j=1
-    while (max(errors)<=2 & j<=loops) {
-
-      norms$mc[[i]][[j]]<-vector("list", 15)
-      norms$ms[[i]][[j]]<-vector("list", 15)
-      if (dists=="Both") {
-        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")  
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
-      } else if (dists=="Normal"){ 
-        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")
-      } else if (dists=="Skew.normal"){
-        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
-      } else {
-        stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew-normal'"))
-      }
-
-      message(paste(i, "Component(s), Loop", j))
-      if (is.na(norms$mc[[i]][[j]][1])==T) {
-        norms$mc[[i]][[j]]<-vector("list")
-        norms$mc[[i]][[j]]$bic<-NA
-        errors[i,1]<-errors[i,1]+1
-      } 
-      if (is.na(norms$ms[[i]][[j]][1])==T) {
-        norms$ms[[i]][[j]]<-vector("list")
-        norms$ms[[i]][[j]]$bic<-NA
-        errors[i,2]<-errors[i,2]+1
-      } 
-      if (dists=="Normal"){ 
-        norms$ms[[i]][[j]]<-vector("list")
-        norms$ms[[i]][[j]]$bic<-NA
-      } else if (dists=="Skew.normal"){
-        norms$mc[[i]][[j]]<-vector("list")
-        norms$mc[[i]][[j]]$bic<-NA
-      }
-      j=j+1
-      ic=i
-    }
-  }
-
-  if (i==maxcp & j==loops+1 & max(errors[maxcp, 2])<=2 & max(errors[maxcp, 1])<=2 ) {
-    nobj<-vector("list")
-    nobj$mc<-matrix(nrow=maxcp, ncol=loops)
-    nobj$ms<-matrix(nrow=maxcp, ncol=loops)
-
-    for (i in 1:maxcp){
-      for (j in 1:loops){
-        nobj$mc[i,j]<-norms$mc[[i]][[j]]$bic
-        nobj$ms[i,j]<-norms$ms[[i]][[j]]$bic
-      }
-    }
-
-    for (i in 1:maxcp){
-      if (dists=="Both" | dists=="Normal"){
-        nobj$lmin$mc[i]<-which.min(nobj$mc[i,])
-      } else {
-        nobj$lmin$mc[i]<-1
-      }
-      if (dists=="Both" | dists=="Skew.normal"){
-        nobj$lmin$ms[i]<-which.min(nobj$ms[i,])
-      } else {
-        nobj$lmin$ms[i]<-1
-      }
-    }
-
-    fitres<-vector("list")
-    for (i in 1:maxcp){
-      fitres$no[[i]]<-vector("list", 15)
-      fitres$sn[[i]]<-vector("list", 15)
-      fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
-      fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
-    }
-
-    nobj<-vector("list")
-
-    for (i in 1:maxcp){
-      for (j in 1:loops){
-        nobj$no[i]<-fitres$no[[i]]$bic
-        nobj$sn[i]<-fitres$sn[[i]]$bic
-      }
-    }
-
-    nobj$minind<-data.frame(no=as.numeric(c(NA)), sn=as.numeric(c(NA)))
-
-    if (is.na(nobj$no[1])==T){
-      nobj$minind$no<-1
-    } else {
-      nobj$minind$no <- (which.min(nobj$no))
-    }
-
-    if (is.na(nobj$sn[1])==T){
-      nobj$minind$sn<-1
-    } else {
-      nobj$minind$sn <- (which.min(nobj$sn))
-    }
-
-    nobj$min<-data.frame(type=c("no", "sn"), bic=as.numeric(c(NA, NA)))
-
-    nobj$min$type<-as.character(c("no", "sn"))
-    type<-c("no", "sn")
-    nobj$min$type<-type
-
-    nobj$min[1,2]<-fitres$no[[nobj$minind$no[1]]]$bic
-    nobj$min[2,2]<-fitres$sn[[nobj$minind$sn[1]]]$bic
-
-    nobj$distnum<-which.min(nobj$min$bic)
-    nobj$dist<-nobj$min$type[nobj$distnum]
-
-    nobj$best<-fitres[[nobj$dist]][[nobj$minind[nobj$dist][1,1]]]
-
-    if (nobj$dist=="no"){
-      nobj$distp="Normal"
-      nobj$distf="Normal"
-    } else if (nobj$dist=="sn") {
-      nobj$distp="Skew-normal"
-      nobj$distf="Skew-normal"
-    } else {
-      nobj$distp=""
-    }
-
-    model<-vector("list")
-
-    model$loops<-loops
-    model$maxcp<-maxcp
-    model$datawithids<-datawithids
-    model$errors<-errors
-    model$loopfits<-fitres
-
-    model$sn2<-fitres$sn[[2]]
-    model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
-    model$norm2<-fitres$no[[2]]
-    model$normbest<-fitres$no[[nobj$minind$no[1]]]
-    model$best<-nobj$best
-    model$bictab<-data.frame(normal=c(NA)[rep(c(1), times=maxcp)], skew.normal=c(NA)[rep(c(1), times=maxcp)])
-    model$bictab$normal<-nobj$no
-    model$bictab$skew.normal<-nobj$sn
-    componentnames<-vector()
-    for (i in 1:maxcp){
-      componentnames[i]<-paste(i, " Component(s):  ", sep="")  
-    }    
-    rownames(model$bictab)<-componentnames
-    colnames(model$bictab)<-c("Normal", "Skew-normal")
-    model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew-normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew-normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
-    rownames(model$summary)<-c("Best Overall", "Best Skew-normal  ", "Best Normal", "Two Skew-normal", "Two Normal")
-    colnames(model$summary)<-c("Description         ", "BIC  ")
-
-    model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
-
-    print("")
-    print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
-    print("")
-    print("Bayesian Information Criterion (BIC) Matrix")
-    print("BIC should be minimized and a difference of 10 BIC indicates strong evidence that the model with lower BIC is superior")
-    print(model$bictab)
-    print("Below is a table of the BIC's of the most common distribution and number of component combinations to base a cutpoint on")
-    print(model$summary)
-
-  return(model)
-
-  } else {
-    if (which.max(errors)<=maxcp){
-      dist<-"Normal"
-    } else {
-      dist<-"Skew.normal"
-    }
-    if (ic<=2){
-      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
-First ensure that the necessary packages have been loaded correctly.
-If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
-    } else {
-      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s).
-Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
-    }
-  }
-
-}
-
-#-----------------------------------------------
-#[IC2]  bicgraph
-#-----------------------------------------------
-#Plots BIC to show which combination of distributions and number of components is optimal by BIC.
-BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior.
-However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.
-
-bicgraph<-function(fitobj,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
-  model<-fitobj
-  color<-setcolor
-  buffer<-(max(model$bictab)-min(model$bictab))/20
-  ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
-  if (is.na(model$bictab$Normal[1])==F) {
-    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of components", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
-    lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
-  } else {
-    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of components", ylab="BIC",  xaxt="n", ylim=ylim)
-    lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
-  }  
-  points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
-  axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
-  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew-normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))
-
-}
-
-#-----------------------------------------------
-#[IC3]  modelpick
-#-----------------------------------------------
-#Selects a specific distribution and format so that it can be read into the uncertainty functions.
-Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.
-The function also performs uncertainty calculations and calculates raw percentages and counts positive and negative for all possible cut points between distributions
-
-modelpick<-function(fitobj, dist="", ncomp=NA){
-  #fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
-  fitres<-fitobj
-  if (dist=="" & is.na(ncomp)==T) {
-    dist=fitres$bestdesc$dist
-    ncomp=fitres$bestdesc$ncomp
-  }
-  if (ncomp==1) {
-    stop("Single component specified so there is no cut-point")
-  }
-
-  data=fitres$datawithids$data
-  n=ncomp
-  if (dist!="Normal" & dist!="Skew-normal") {
-    stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
-  } else if ((dist=="Normal" & is.na(fitres$loopfits$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$loopfits$sn[[1]]$bic)==T) | ncomp>length(fitres$loopfits$no) ){
-    stop("Must choose a distribution and number of components that were fitted in the fitloops function")
-  } else {
-    if (dist=="Normal"){
-      dist="no"
-    } else {
-      dist="sn"
-    }
-
-    if (dist=="no"){
-      distp="Normal"
-    } else if (dist=="sn") {
-      distp="Skew-Normal"
-    } else {
-      distp=""
-    }
-
-    singlefit<-vector("list")
-    singlefit<-fitres$loopfits[[dist]][[ncomp]]
-
-    comp<-data.frame(ogroup=singlefit$group, data=data)
-    comp$index <- as.numeric(row.names(comp))
-    comp <- comp[order(comp$data),] 
-    comp$diff<-c(0, diff(comp$ogroup))
-    nl=as.numeric(dim(comp)[1])
-
-    #Will recode the distributions so that they are in order with 1 being the lowest
-    comp$c<-vector(mode="numeric", length=nl)
-    comp$c[1]<-1
-    for (i in 2:nl){
-      if (comp$diff[i]==0) {
-        comp$c[i]=comp$c[i-1]
-      } else {
-        comp$c[i]=comp$c[i-1]+1
-      }
-    }
-    comp <- comp[order(comp$index),]
-    singlefit$group2<-comp$c
-    comp <- comp[order(comp$data),]
-
-    intDistErr = function() {
-      tryCatch(data.frame(c=unique(comp$c), o=unique(comp$ogroup)),
-               error = function(e) {data.frame(c=c(NA)[rep(c(1), times=length(comp$c))], o=c(NA)[rep(c(1), times=length(comp$c))])}) 
-    }
-
-    cross<-intDistErr()
-    if (suppressWarnings(is.na(cross[1,1])==T)){
-      stop("This combination of distribution and number of components yeilds interrupting distributions, which make determining a cut-point impossible.
-Please choose a different combination of distribution and number of components")
-    }
-
-    for (i in 1:n){
-      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
-      singlefit$sigma2[i]<-fitres$loopfits[[dist]][[ncomp]]$sigma2[cross$o[i]]
-      singlefit$shape[i]<-fitres$loopfits[[dist]][[ncomp]]$shape[cross$o[i]]
-      singlefit$pii[i]<-fitres$loopfits[[dist]][[ncomp]]$pii[cross$o[i]]
-      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
-    }
-
-    singlefit$obs.prob<-NULL
-
-    #Detect if there are any interupting distributions (can also be recoded to kick someone out and tell them to pick something else:
-    if (max(comp$c)>max(comp$ogroup)) {
-      print("Warning: there is at least one interupting distribution, which may make the determination of a cutoff and the calculation of an indeterminate range difficult in later functions.
-If later functions fail, try a different combination of distribution and number of components (ideally the next most optimal by BIC).")
-    }
-
-    par<-vector("list")
-    par<-data.frame(mean=c(NA)[rep(c(1), times=n)], sd=c(NA)[rep(c(1), times=n)], shape=c(NA)[rep(c(1), times=n)], pii=c(NA)[rep(c(1), times=n)])
-    for (i in 1:n){
-      par$mean[i]=singlefit$mu[i]
-      par$sd[i]=sqrt(singlefit$sigma2[i])
-      par$sigma2[i]=singlefit$sigma2[i]
-      par$shape[i]=singlefit$shape[i]
-      par$pii[i]=singlefit$pii[i]
-    }
-  }
-
-  #Performs uncertainty calculations.
-  #uncert<-function(fitpickobj){
-
-  fit<-singlefit
-  #Create Uncertainties
-  v<-c(seq(min(data),max(data),0.0001))
-
-  nl<-length(v)
-  PD<-vector(mode="numeric", length=n)
-  PD[1]<-1
-  g1<-0
-  g2<-0
-  Pv.D<-matrix(0, nl, n)
-  Pv.DPD<-matrix(0, nl, n)
-  PDT<-matrix(0, nl, 1)
-  PD.v<-matrix(0, nl, n)
-  for (i in 2:n) {
-    PD[i]<-par$pii[i]
-    PD[1]<-PD[1]-PD[i]
-  }
-  for (i in 1:n){
-    if (dist=="Normal"){
-      Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
-    } else {
-      Pv.D[,i]<-dsn(v,xi=par$mean[i],omega=par$sd[i],alpha=par$shape[i])
-    }
-  }
-  for (i in 1:n){
-    for (j in 1:nl){
-      Pv.DPD[j,i]<-Pv.D[j,i]*PD[i]
-    }
-  }
-  PDT<-rowSums (Pv.DPD)
-  for (i in 1:n){
-    for (j in 1:nl){
-      PD.v[j,i] <- Pv.DPD[j,i] / (PDT[j])
-    }
-  }
-
-  mcluncert<-1-apply(PD.v, 1, max) 
-  uncertmat<-PD.v
-  restrmat<-cbind(PD.v, v)
-  uncertout<-vector("list")
-
-  #multcut<-function(uncertobj){
-
-  closest<-function(uncertdf,min,max,level){
-    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
-    diff.cut<-abs(uncertdf2$uncertainty.v - level)
-    min.diffc<-min(diff.cut)
-    y<-which(min.diffc==diff.cut)
-    #take first one
-    z<-min(y)
-    cut<-uncertdf2$v[z]  
-    return(cut) 
-  }
-  g1<-vector("list")
-  g2<-vector("list")
-  uncertdf<-vector("list")
-  uncertainty.v<-vector("list")
-  cutpoint<-c(0)[rep(c(1), times=ncomp-1)]
-  for (j in 1:(ncomp-1)){
-    g1[[j]]<-0
-    g2[[j]]<-0
-    for (i in 1:j){
-      g1[[j]]<-g1[[j]]+PD.v[,i]
-    }
-    for (i in (j+1):n){
-      g2[[j]]<-g2[[j]]+PD.v[,i]
-    }
-    uncertainty.v[[j]]<-1-pmax(g1[[j]], g2[[j]])
-  }
-
-  for (j in 1:(ncomp-1)){
-    uncertdf[[j]]<-data.frame(v, uncertainty.v=uncertainty.v[[j]])
-    cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
-  }
-  class2<-vector("list")
-  for (j in 1:(ncomp-1)){
-    class2[[j]]<-ifelse(data>=cutpoint[j], "positive", "negative")
-  }
-  class<- data.frame(matrix(unlist(class2), nrow=singlefit$n, byrow=F),stringsAsFactors=TRUE)
-  class$data<-data
-
-  table<-vector("list", ncomp-1)
-  for (i in 1:(ncomp-1)){
-    table[[i]]<-table(class2[[i]])
-
-  }
-  npos<-vector("numeric")
-  nneg<-vector("numeric")
-  dpos<-vector("numeric")
-  dneg<-vector("numeric")
-  ppos<-vector("numeric")
-  pneg<-vector("numeric")
-  grpname<-vector("numeric")
-  for (i in 1:(ncomp-1)){
-    npos[i]<-table[[i]][2]
-    nneg[i]<-table[[i]][1]
-    dpos[i]<-npos[i]/(npos[i]+nneg[i])
-    dneg[i]<-1-dpos[i]
-    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
-    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
-    grpname[i]<-paste("Cut between components ", i, " and ", i+1, sep="")
-  }
-  classtab<-data.frame( nneg, npos, dneg, dpos)
-  outtab<-data.frame(nneg, npos, pneg, ppos)
-  rownames(outtab) <- grpname
-  colnames(outtab) <-c("Number Negative", "Number Positive", "Percent Negative", "Percent Positive")
-
-  cutobj<-vector("list")
-
-  cutobj$singlefit<-singlefit
-  cutobj$par<-par
-  cutobj$datawithids<-fitres$datawithids
-  cutobj$desc<-vector("list")
-  cutobj$desc$ncomp<-ncomp
-  cutobj$desc$dist<-distp
-
-  cutobj$uncertmat<-uncertmat
-  cutobj$mcluncert<-mcluncert
-  cutobj$restrmat<-restrmat
-  cutobj$v<-v
-
-  cutobj$cutpoint<-cutpoint
-  cutobj$class<-class
-  cutobj$uncertdf<-uncertdf
-  cutobj$type<-"Standard"
-  cutobj$g1<-g1
-  cutobj$g2<-g2
-  cutobj$classtab<-classtab
-  cutobj$outtab<-outtab
-  cutobj$cutnames<-grpname
-
-  print(outtab)
-
-  return(cutobj)
-}
-
-#-----------------------------------------------
-#[IC4]  rawuncertgraph
-#-----------------------------------------------
-#Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid.
-Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not.
-In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
-  "green4",          #1
-  "turquoise3",      #2
-  "royalblue2",      #3
-  "darkorchid2",     #4
-  "firebrick3",      #5
-  "darkorange2",     #6
-  "white",           #7
-  "grey",            #8
-  "black"            #9
-)){
-
-  grobj<-list()
-  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
-    grobj$maxval<-max(modelpickobj$datawithids$data)
-  } else {
-    grobj$minval<-xlim[1]
-    grobj$maxval<-xlim[2]
-  }
-
-  plot(modelpickobj$v, modelpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")
-
-  col<-setcolor
-  ncomp<-modelpickobj$desc$ncomp
-  for (i in 1:(modelpickobj$desc$ncomp-1)){
-    abline(v =modelpickobj$cutpoint[i], untf = T, col=col[i], lwd=2)
-  }
-
-  if (suppresslegend==F) {
-    legtxt<-vector("character", ncomp)
-    colors<-vector("character", ncomp)
-    for (i in 2:ncomp){
-      legtxt[i]<-paste(modelpickobj$cutnames[i-1], " , x=", round(modelpickobj$cutpoint[i-1], 2), sep="")
-      colors[i]<-col[i-1]
-    }  
-    legtxt[1]<-"Uncertainty"
-    colors[1]<-col[length(col)]
-    linewidth<-c(2)[rep(c(1), times=ncomp)]
-    legend("topright", col=colors, lwd=c(1,linewidth), legend=legtxt, xjust=1, seg.len=3)
-  }
-}
-
-#-----------------------------------------------
-#[IC5]  rawdistgraph
-#-----------------------------------------------
-#Displays the distributions of all the components for the chosen combination of distribution and number of components. 
-rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
-  "green4",          #1
-  "turquoise3",      #2
-  "royalblue2",      #3
-  "darkorchid2",     #4
-  "firebrick3",      #5
-  "darkorange2",     #6
-  "white",           #7
-  "grey",            #8
-  "black"            #9
-)){
-  #Parameter specification
-  ncomp=modelpickobj$desc$ncomp
-  grobj<-vector("list")
-  grobj$gpar<-vector("list")
-  data<-modelpickobj$datawithids$data
-  color<-setcolor
-
-  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
-    grobj$maxval<-max(data)
-  } else {
-    grobj$minval<-xlim[1]
-    grobj$maxval<-xlim[2]
-  }
-
-  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)
-
-  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
-  grobj$gt<-vector("list", ncomp+1)
-  grobj$maxdens<-vector("numeric", ncomp)
-
-  grobj$coloropt<-vector("list", 5)
-  grobj$coloropt[[1]]<-c(color[1], color[9])
-  grobj$coloropt[[2]]<-c(color[1], color[5], color[9])
-  grobj$coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
-  grobj$coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
-  grobj$coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
-  grobj$colors<-grobj$coloropt[[ncomp]]
-
-  for (i in 1:ncomp){
-    grobj$gpar$mean[i]=modelpickobj$par$mean[i]
-    grobj$gpar$sd[i]=sqrt(modelpickobj$par$sigma2[i])
-    grobj$gpar$shape[i]=modelpickobj$par$shape[i]
-    grobj$gpar$pii[i]=modelpickobj$par$pii[i]
-  }
-  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
-  for (i in 1:ncomp) {
-    grobj$gt[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
-  }
-  grobj$gt[[ncomp+1]]<-vector("numeric", 1000)
-  for (i in 1:ncomp) {
-    grobj$gt[[ncomp+1]]<-grobj$gt[[ncomp+1]]+grobj$gt[[i]]
-  }
-  grobj$ylim<-c(0, max(grobj$gt[[ncomp+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
-  grobj$xlim<-c(grobj$minval,grobj$maxval)
-
-  dist<-c(0)[rep(c(0), times=ncomp)]
-  for (i in 1:ncomp){
-    if (modelpickobj$singlefit$shape[i]>0) {
-      dist[i]<-1
-    }
-  }
-
-  #Plot Graph
-
-  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, modelpickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
-  for (i in 1:ncomp) {
-    lines(grobj$t, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
-  }
-  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
-  legtxt<-vector("character", ncomp+1)
-  colors<-vector("character", ncomp+1)
-  for (i in 1:ncomp){
-    legtxt[i]<-paste("Distribution of Component ", i, sep="")
-    colors[i]<-grobj$colors[i]
-  }  
-  legtxt[ncomp+1]<-"Overall Distribution"
-  colors[ncomp+1]<-grobj$colors[length(grobj$colors)]
-  linetype<-c(1)[rep(c(1), times=ncomp)]
-  if (suppresslegend==F) {
-    legend("topright", col=colors, lwd=2, legend=legtxt, xjust=1, seg.len=3, title="Distributions", lty=c(linetype,2))
-  }
-
-  #return(grobj)
-}
-
-#-----------------------------------------------
-#[IC6]  rawhistcuts
-#-----------------------------------------------
-#Displays which component each bar of the histogram would be assigned to.
-
-rawhistcuts<-function(modelpickobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
-  "green4",          #1
-  "turquoise3",      #2
-  "royalblue2",      #3
-  "darkorchid2",     #4
-  "firebrick3",      #5
-  "darkorange2",     #6
-  "white",           #7
-  "grey",            #8
-  "black"            #9
-)){
-  data<-modelpickobj$datawithids$data
-  grobj<-list()
-
-  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    grobj$minval<-(1/10)*trunc(10*min(data)) 
-    grobj$maxval<-max(data)
-    hist<-hist(data, breaks=setbreaks, plot=F)
-  } else {
-    grobj$minval<-xlim[1]
-    grobj$maxval<-xlim[2]
-    datalim<-data[ which(data>=grobj$minval & data<=grobj$maxval)]
-    hist<-hist(datalim, breaks=setbreaks, plot=F)
-  }
-
-  ncomp<-modelpickobj$desc$ncomp
-  cuts <- cut(hist$breaks, c(-Inf,modelpickobj$cutpoint,Inf))
-
-  color<-setcolor
-  coloropt<-vector("list", 5)
-  coloropt[[1]]<-c(color[1], color[9])
-  coloropt[[2]]<-c(color[1], color[5], color[9])
-  coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
-  coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
-  coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
-  col<-coloropt[[ncomp]]
-
-  cuts<-as.numeric(cuts)
-  colors<-vector("character", length(cuts))
-  for (i in 1:max(cuts)){
-    for (j in 1:length(cuts)){
-      if (cuts[j]==i) {
-        colors[j]<-col[i]
-      }
-    }
-  }
-  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", modelpickobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")
-
-  if (suppresslegend==F){
-    legtxt<-vector("character", ncomp)
-    colors<-vector("character", ncomp)
-    for (i in 1:ncomp){
-      legtxt[i]<-paste("Component ", i, sep="")
-      colors[i]<-col[i]
-    }  
-    linetype<-c(1)[rep(c(1), times=ncomp)]
-    legend("topright", col=colors, lwd=3, legend=legtxt, xjust=1, seg.len=3, title="Components", lty=linetype)
-  }
-}
-
-#-----------------------------------------------
-#[IC7]  cutoff
-#-----------------------------------------------
-#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (80 and 90), as well as classification into negative, indeterminate and positive components.
-It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate.
-In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
-cutoff<-function(modelpickobj, cutcomp=0, standardcert=T, newcertlevel=0){
-  data<-modelpickobj$datawithids$data
-  n=modelpickobj$desc$ncomp
-  certlevel=newcertlevel
-  if (n==2){
-    cutcomp=1
-  } else if (cutcomp==0) {
-    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
-  }
-  if ((certlevel>=1 & certlevel!=0) | (certlevel<0.5 & certlevel!=0)){
-    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 80% set certlevel=0.80")
-  }
-  if (standardcert==F & newcertlevel==0){
-    stop("Neither standard nor non-standard certainty levels are requested.
-For standard certainty levels set standardcert=T.
-For non-standard certainty levels set newcertlevel to desired certainty.
-Both standard and nonstandard certainties can be generated in the same function call.")
-  }
-  uncertdf<-modelpickobj$uncertdf[[cutcomp]]
-  closest<-function(uncertdf,min,max,level){
-    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
-    diff.cut<-abs(uncertdf2$uncertainty.v - level)
-    min.diffc<-min(diff.cut)
-    y<-which(min.diffc==diff.cut)
-    #take first one
-    z<-min(y)
-    cut<-uncertdf2$v[z]  
-    return(cut) 
-  }
-  crosses<-function(uncertdf,min,max,level){
-    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
-    diff.cut<-abs(uncertdf2$uncertainty.v - level)
-    min.diffc<-min(diff.cut)
-    if (min.diffc<=0.001){
-      y<-which(min.diffc==diff.cut)
-      #take first one
-      z<-min(y)
-      cut<-uncertdf2$v[z]  
-    } else {
-      cut<-NA
-    }
-    return(cut)
-  }
-
-  cutpoint<-modelpickobj$cutpoint[cutcomp]  
-  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
-  minuncub<-closest(uncertdf,cutpoint,max(data),0)
-  class<-data.frame(id=modelpickobj$datawithids$id, data=modelpickobj$datawithids$data)
-  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
-  class$groupdi<-as.factor(class$groupdi)
-
-  if (standardcert==T){
-    lb80<-crosses(uncertdf,minunclb,cutpoint,0.2)  
-    ub80<-crosses(uncertdf,cutpoint,minuncub,0.2)
-    lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
-    ub90<-crosses(uncertdf,cutpoint,minuncub,0.1)
-    bound80<-c(lb80, ub80)
-    bound90<-c(lb90, ub90)
-    if (is.na(lb80)==T){
-      message("Lower bound of the 80% indeterminate range not found because uncertainty is never below 20% to the left of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
-    }
-    if (is.na(ub80)==T){
-      message("Upper bound of the 80% indeterminate range not found because uncertainty is never below 20% to the right of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
-    }
-    if (is.na(lb90)==T){
-      message("Lower bound of the 90% indeterminate range not found because uncertainty is never below 10% to the left of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
-    }
-    if (is.na(ub90)==T){
-      message("Upper bound of the 90% indeterminate range not found because uncertainty is never below 10% to the right of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.")
-    }
-    class$group90<-"indeterminate"
-    class$group90[class$data<=bound90[1]]<-"negative"
-    class$group90[class$data>=bound90[2]]<-"positive"
-    class$group80<-"indeterminate"
-    class$group80[class$data<=bound80[1]]<-"negative"
-    class$group80[class$data>=bound80[2]]<-"positive"
-    class$group90<-as.factor(class$group90)
-    class$group80<-as.factor(class$group80)
-    class$group80 = factor(class$group80,levels(class$group80)[c(2,3,1)])
-    class$group90 = factor(class$group90,levels(class$group90)[c(2,3,1)])
-    class$group<-NA
-  }
-
-  if (newcertlevel>0){
-    uncertlevel<-1-certlevel
-    uncertlevelpct<-certlevel
-    lbuns<-crosses(uncertdf,minunclb,cutpoint,uncertlevel)  
-    ubuns<-crosses(uncertdf,cutpoint,minuncub,uncertlevel)
-    bound<-c(lbuns, ubuns)
-    if (is.na(lbuns)==T){
-      message(paste("Lower bound of the ", 100*certlevel, "% indeterminate range not found because uncertainty is never below ", 100*(1-certlevel), "% to the left of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.", sep=""))
-    }
-    if (is.na(ubuns)==T){
-      message(paste("Upper bound of the ", 100*certlevel, "% indeterminate range not found because uncertainty is never below ", 100*(1-certlevel), "% to the right of the cutpoint.
-Consider adjusting the non-standard certainty option (newcertlevel) to find a more descriptive indeterminate range.", sep=""))
-    }
-    if (standardcert==F){
-      class$group90<-NA
-      class$group80<-NA
-    }
-    class$group<-"indeterminate"
-    class$group[class$data<=bound[1]]<-"negative"
-    class$group[class$data>=bound[2]]<-"positive"
-    class$group<-as.factor(class$group)
-    class$group = factor(class$group,levels(class$group)[c(2,3,1)])
-  }
-
-  cutobj<-vector("list")
-  cutobj$cutpoint<-cutpoint
-  cutobj$bound80<-c(NA,NA)
-  cutobj$bound90<-c(NA,NA)
-  cutobj$bound<-c(NA,NA)
-  cutobj$class<-class
-  cutobj$uncertdf<-uncertdf
-
-  if (standardcert==T & newcertlevel==0){
-    cutobj$type<-"Standard"
-    cutobj$bound80<-bound80
-    cutobj$bound90<-bound90
-  } else if (standardcert==F & newcertlevel>0) {
-    cutobj$type<-"Non.Standard"
-    cutobj$bound<-bound
-  } else {
-    cutobj$type<-"Standard.and.Non"
-    cutobj$bound80<-bound80
-    cutobj$bound90<-bound90
-    cutobj$bound<-bound
-  }
-
-  cutobj$par<-modelpickobj$par
-  cutobj$uncertobj<-modelpickobj$uncertmat
-  cutobj$density<-data.frame(g1=modelpickobj$g1[[cutcomp]], g2=modelpickobj$g2[[cutcomp]])
-  cutobj$desc<-modelpickobj$desc
-  cutobj$desc$cutcomp<-cutcomp
-  cutobj$desc$certlevel<-newcertlevel
-  cutobj$datawithids<-modelpickobj$datawithids
-
-  return(cutobj) 
-}
-
-#-----------------------------------------------
-#[IC8] cutuncertgraph
-#-----------------------------------------------
-#Displays the uncertainty function after components have been combined to create positive and negative components with cut-points and bounds of indeterminate range(s) overlaid.
-Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
-cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
-  "green4",          #1
-  "turquoise3",      #2
-  "royalblue2",      #3
-  "darkorchid2",     #4
-  "firebrick3",      #5
-  "darkorange2",     #6
-  "white",           #7
-  "grey",            #8
-  "black"            #9
-)){
-  color=setcolor
-  grobj<-vector("list")
-
-  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
-    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty")
-  } else {
-    grobj$minval<-xlim[1]
-    grobj$maxval<-xlim[2]
-    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty", xlim=c(grobj$minval,grobj$maxval))
-  }
-
-  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
-  if (cutobj$type=="Standard") {
-    abline(h =0.1, untf = T, col=color[4], lwd=1, lty=2)
-    abline(h =0.2, untf = T, col=color[2], lwd=1, lty=2)
-    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
-    legtxt<-c("Uncertainty",
-              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-              "Uncertainty=0.20",
-              "Indeterminate Range with", 
-              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
-              "Uncertainty=0.10",
-              "Indeterminate Range with", 
-              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep ="")
-    )
-    if (suppresslegend==F){
-      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
-      }
-  } else if (cutobj$type=="Non.Standard") {
-    abline(h =(1-cutobj$desc$certlevel), untf = T, col=color[3], lwd=1, lty=2)
-    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
-    legtxt<-c("Uncertainty",
-              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-              paste("Uncertainty=", (1-cutobj$desc$certlevel), sep =""),
-              "Indeterminate Range with", 
-              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
-    )
-    if (suppresslegend==F){
-      legend("topright", col=c(color[9],color[6], color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1), legend=legtxt, xjust=1, seg.len=1, title="Cutpoint, Uncertainties, and Boundaries")
-    }
-  } else if (cutobj$type=="Standard.and.Non") {
-    abline(h =0.1, untf = T, col=color[4], lwd=1, lty=2)
-    abline(h =0.2, untf = T, col=color[2], lwd=1, lty=2)
-    abline(h =(1-cutobj$desc$certlevel), untf = T, col=color[3], lwd=1, lty=2)
-    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
-    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
-    legtxt<-c("Uncertainty",
-              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-              "Uncertainty=0.20",
-              "Indeterminate Range with", 
-              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
-              "Uncertainty=0.10",
-              "Indeterminate Range with", 
-              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
-              paste("Uncertainty=", (1-cutobj$desc$uncertlevel), sep =""),
-              "Indeterminate Range with", 
-              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
-    )
-    if (suppresslegend==F){
-      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white", color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
-    }
-  }
-}
-
-#-----------------------------------------------
-#[IC9]  cutdistgraph
-#-----------------------------------------------
-#Displays the distributions of the positive and negative components with cut-points and bounds of indeterminate range(s) overlaid.
-Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
-cutdistgraph<-function(cutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
-  "green4",          #1
-  "turquoise3",      #2
-  "royalblue2",      #3
-  "darkorchid2",     #4
-  "firebrick3",      #5
-  "darkorange2",     #6
-  "white",           #7
-  "grey",            #8
-  "black"            #9
-)){
-  #Parameter specification
-  ncomp<-cutobj$desc$ncomp
-  cutcomp<-cutobj$desc$cutcomp
-  cutobj<-cutobj
-  data<-cutobj$datawithids$data
-  color<-setcolor
-
-  grobj<-vector("list")
-  grobj$gpar<-vector("list")
-
-  if (is.na(xlim[1])==T | is.na(xlim[2])==T) { 
-    grobj$minval<-(1/10)*trunc(10*min(data))     
-    grobj$maxval<-max(data)
-  } else {
-    grobj$minval<-xlim[1]
-    grobj$maxval<-xlim[2]
-  }
-
-  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)
-
-  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
-  grobj$gs<-vector("list", ncomp+1)
-  grobj$gt<-vector("list", 2+1)
-  grobj$maxdens<-vector("numeric", ncomp)
-  grobj$colors<-c(setcolor[1], setcolor[5], setcolor[9])
-  grobj$t<-cutobj$uncertdf$v
-
-  for (i in 1:ncomp){
-    grobj$gpar$mean[i]=cutobj$par$mean[i]
-    grobj$gpar$sd[i]=sqrt(cutobj$par$sigma2[i])
-    grobj$gpar$shape[i]=cutobj$par$shape[i]
-    grobj$gpar$pii[i]=cutobj$par$pii[i]
-  }
-  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
-  for (i in 1:ncomp) {
-    grobj$gs[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
-  }
-  grobj$gt[[1]]<-0
-  grobj$gt[[2]]<-0
-  grobj$gt[[3]]<-0
-  for (i in 1:cutcomp){
-    grobj$gt[[1]]<-grobj$gt[[1]]+grobj$gs[[i]]
-  }
-  for (i in (cutcomp+1):ncomp){
-    grobj$gt[[2]]<-grobj$gt[[2]]+grobj$gs[[i]]
-  }
-  for (i in 1:ncomp) {
-    grobj$gt[[3]]<-grobj$gt[[3]]+grobj$gs[[i]]
-  }
-
-  grobj$ylim<-c(0, max(grobj$gt[[2+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
-  grobj$xlim<-c(round(min(data),digits=1),grobj$maxval)
-
-  #Plot Graph
-  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste(cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "), border=color[8], ylim=grobj$ylim, xlim=grobj$xlim)
-
-  for (i in 1:2) {
-    lines(cutobj$uncertdf$v, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
-  }
-  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
-  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
-  if (cutobj$type=="Standard") {
-    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
-    if (suppresslegend==F) {
-      legtxt<-c("Negative Distribution",
-                "Positive Distribution",
-                "Overall Distribution",
-                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-                "Indeterminate Range with", 
-                paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
-                "Indeterminate Range with", 
-                paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep ="")
-      )
-      legend("topright", col=c(color[1], color[5], color[9], color[6], color[2], "white", color[4], "white"), lwd=c(2,2,2,1,1,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1,1,1))
-    }
-  } else if (cutobj$type=="Non.Standard") {
-    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
-    if (suppresslegend==F){
-      legtxt<-c("Negative Distribution",
-                "Positive Distribution",
-                "Overall Distribution",
-                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-                "Indeterminate Range with", 
-                paste((cutobj$desc$uncertlevel*100), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
-      )
-      legend("topright", col=c(color[1], color[5], color[9], color[6], color[3], "white"), lwd=c(2,2,2,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1))
-    }
-  } else {
-    abline(v =cutobj$bound80, untf = T, col=color[2], lwd=1)
-    abline(v =cutobj$bound90, untf = T, col=color[4], lwd=1)
-    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
-    if (suppresslegend==F){
-    legtxt<-c("Negative Distribution",
-              "Positive Distribution",
-              "Overall Distribution",
-              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
-              "Indeterminate Range with", 
-              paste("80% Certainty: (", round(cutobj$bound80[1], 2), " , ", round(cutobj$bound80[2], 2), ")", sep =""),
-              "Indeterminate Range with", 
-              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
-              "Indeterminate Range with", 
-              paste((100*cutobj$desc$certlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
-      )
-    legend("topright", col=c(color[1], color[5], color[9], color[6], color[2], "white", color[4], "white", color[3], "white"), lwd=c(2,2,2,1,1,1,1,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1,1,1,1,1))
-    }
-  }
-
-  #return(grobj)
-
-}
-
-#-----------------------------------------------
-#[IC10]  summaryout
-#-----------------------------------------------
-
-#Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 80 and 90 indeterminate ranges and (if desired) one non-standard uncertainty level.
-summaryout<-function(cutobj,fileandpathname=NULL){
-
-  if (is.null(cutobj)==T) {
-    print("Cutoff object must be specified")
-  } 
-  if (cutobj$type=="Standard"){
-    dataout<-cutobj$class[c("id", "data", "group80", "group90","groupdi")]
-    names(dataout)<-c(names(cutobj$class[c(1,2,5,4,3)]))
-  } else if (cutobj$type=="Non.Standard"){
-    dataout<-cutobj$class[c("id", "data", "group", "groupdi")]
-    names(dataout)<-c(names(cutobj$class[c(1,2)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
-  } else {
-    dataout<-cutobj$class[c("id", "data","group80", "group90", "group", "groupdi")]
-    names(dataout)<-c(names(cutobj$class[c(1,2,5,4)]),paste('group', (100*cutobj$desc$certlevel), sep=''), names(cutobj$class[3]))
-  }
-
-  if (is.null(fileandpathname)==F){
-    write.csv(dataout, file =paste(fileandpathname, ".csv", sep=""), row.names=F)
-  }
-
-  outdataobj<-dataout
-
-  comptable <- function(table){
-
-    dftable<-as.data.frame(table)
-    len<-nrow(dftable)
-    dftable$orders<-c(NA)[rep(c(1), times=len)]
-    for (i in 1:len){
-      if (as.character(dftable$Var1[i])=="negative"){
-        dftable$orders[i]<-0
-      } else if (as.character(dftable$Var1[i])=="indeterminate"){
-        dftable$orders[i]<-1
-      } else if (as.character(dftable$Var1[i])=="positive"){
-        dftable$orders[i]<-2
-      }
-    }
-    row.names(dftable)<-dftable$Var1
-    dftable$Var1<-NULL
-    if (is.na(dftable["indeterminate", 1])==T) {
-      Freq<-c(dftable$Freq,0)
-      grps2<-as.data.frame(Freq)
-      x<-c(row.names(dftable),"indeterminate")
-      row.names(grps2)<-x
-      grps2$orders<-c(dftable$orders,1)
-      dftable<-grps2
-      Freq<-NULL
-      x<-NULL
-    }
-    if (is.na(dftable["negative", 1])==T) {
-      Freq<-c(dftable$Freq,0)
-      grps2<-as.data.frame(Freq)
-      x<-c(row.names(dftable),"negative")
-      row.names(grps2)<-x
-      grps2$orders<-c(dftable$orders,0)
-      dftable<-grps2
-      Freq<-NULL
-      x<-NULL
-    }
-    if (is.na(dftable["positive", 1])==T) {
-      Freq<-c(dftable$Freq,0)
-      grps2<-as.data.frame(Freq)
-      x<-c(row.names(dftable),"positive")
-      row.names(grps2)<-x
-      grps2$orders<-c(dftable$orders,2)
-      dftable<-grps2    
-      Freq<-NULL
-      x<-NULL
-    }
-    dftable<-dftable[order(dftable$orders),]
-    return(dftable)
-  }
-
-  grps<-table(outdataobj$groupdi)
-  grps<-comptable(grps)
-  grps$group<-row.names(grps)
-  grps$groupdi<-grps$Freq
-  grps$Freq<-NULL
-
-  if (ncol(outdataobj)!=5 ){
-    grpsns<-table(outdataobj[,4])
-    grpsns<-comptable(grpsns)
-    grps$gns<-grpsns$Freq
-    names(grps)<-c(names(grps[1:3]),paste(names(outdataobj)[4]))
-  }
-
-  if (ncol(outdataobj)!=4 ){ 
-    grps80<-table(outdataobj$group80)
-    grps80<-comptable(grps80)
-    grps$group80<-grps80$Freq
-    grps90<-table(outdataobj$group90)
-    grps90<-comptable(grps90)
-    grps$group90<-grps90$Freq
-  }
-
-  names(grps)
-
-  for (i in names(grps)){
-    nneg<-vector("numeric")
-    nind<-vector("numeric")
-    npos<-vector("numeric")
-
-    dneg<-vector("numeric")  
-    dind<-vector("numeric")
-    dpos<-vector("numeric")
-
-    pneg<-vector("numeric")
-    pind<-vector("numeric")
-    ppos<-vector("numeric")
-  }
-
-  vars<-c("groupdi", "group80", "group90", "group80")
-
-  grpname<-vector("numeric")
-  for (i in 1:(ncol(grps)-2)){
-    nneg[i]<-grps[1,i+2]
-    nind[i]<-grps[2,i+2] 
-    npos[i]<-grps[3,i+2]
-  }
-  for (i in 1:(ncol(grps)-2)){
-    dneg[i]<-nneg[i]/(npos[i]+nind[i]+nneg[i])
-    dind[i]<-nind[i]/(npos[i]+nind[i]+nneg[i])
-    dpos[i]<-npos[i]/(npos[i]+nind[i]+nneg[i])
-    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
-    pind[i]<-paste(round(100*dind[i], 2), "%", sep="")
-    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
-  }
-  if (ncol(grps)==4){
-    grpname<-c("Raw Cutoff", paste("Cutoff with ", cutobj$desc$certlevel*100, "% Certainty", sep=""))
-  } else if (ncol(grps)==5) {
-    grpname<-c("Raw Cutoff", "Cutoff with 80% Certainty", "Cutoff with 90% Certainty")
-  } else if (ncol(grps)==6) {
-    grpname<-c("Raw Cutoff", paste("Cutoff with ", cutobj$desc$certlevel*100, "% Certainty", sep=""), "Cutoff with 80% Certainty", "Cutoff with 90% Certainty")
-  }
-
-  classtab<-data.frame(nneg, nind, npos, dneg, dind, dpos)
-  outtab<-data.frame(nneg, nind, npos, pneg, pind, ppos)
-  rownames(outtab) <- grpname
-  colnames(outtab) <-c("No. Neg", "No. Indet", "No. Pos", "% Neg", "% Indet", "% Pos")
-
-  model<-vector("list")
-  if (is.null(cutobj)==F){
-    model$cutpoint<-cutobj$cutpoint
-    model$desc<-cutobj$desc
-    model$bounds<-vector("list")
-    model$bounds$boundns<-cutobj$bound
-  } else {
-    model$cutpoint<-cutobj$cutpoint     
-    model$desc<-cutobj$desc
-    model$bounds<-vector("list")
-    model$bounds$bound80<-cutobj$bound80
-    model$bounds$bound90<-cutobj$bound90
-  }
-  if (is.null(cutobj)==F & is.null(cutobj)==F){
-    model$bounds$bound80<-cutobj$bound80
-    model$bounds$bound90<-cutobj$bound90
-  }
-  model$classtab<-classtab
-  model$outtab<-outtab
-  model$classification<-outdataobj
-
-  message(paste("The absolute cutoff is", round(cutobj$cutpoint,3), sep=" "))
-  message("Boundries of Indeterminates:")
-  if (cutobj$type!="Standard"){
-    message(paste(cutobj$desc$certlevel*100, "% Certainty Range: (", round(cutobj$bound[1], 3), ", ", round(cutobj$bound[2], 3), ")", sep=""))
-  }
-  if (cutobj$type!="Non.Standard"){  
-    message(paste("80% Certainty Range: (", round(cutobj$bound80[1], 3), ", ", round(cutobj$bound80[2], 3), ")", sep=""))
-    message(paste("90% Certainty Range: (", round(cutobj$bound90[1], 3), ", ", round(cutobj$bound90[2], 3), ")", sep=""))
-  }
-
-  return(model)
-}
-
+source("./Cutoff_Algorithm.R")
 ```
@@
 
-hist(examp1$data, breaks=100, main="Histogram of Data")
+hist(examp1$data, breaks=100, main="", xlab="Optical Density")
 dens1<-density(examp1$data)
-plot(dens1, main="Density of Data")
+plot(dens1, main="", xlab="Optical Density")
 
@@
 
-rawuncertgraph(modelpickobject1)
-rawdistgraph(modelpickobject1)
-rawhistcuts(modelpickobject1)
+rawuncertgraph(modelpickobject1, title="", xlab="Optical Density")
+rawdistgraph(modelpickobject1, title="", xlab="Optical Density")
+rawhistcuts(modelpickobject1, title="", xlab="Optical Density")
 
-cutoffobject1<-cutoff(modelpickobj = modelpickobject1)
+cutoffobject1<-cutoff(modelpickobj = modelpickobject1, newcertlevel = 0.85)
 
-cutuncertgraph(cutobj = cutoffobject1)
-cutdistgraph(cutobj = cutoffobject1)
+cutuncertgraph(cutobj = cutoffobject1, title="", xlab="Optical Density")
+cutdistgraph(cutobj = cutoffobject1, title="", xlab="Optical Density")
 
@@
 examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", stringsAsFactors = F, colClasses=c("NULL", NA, NA))
-examp2<-examp2[,c("id", "data")]
 
-hist(examp2$data, breaks=100, main="Histogram of Data")
+hist(examp2$data, breaks=100, main="", xlab="Optical Density")
 dens2<-density(examp2$data)
-plot(dens2, main="Density of Data")
+plot(dens2, main="", xlab="Optical Density")
 
@@
 
-rawuncertgraph(modelpickobj = modelpickobject2)
-rawdistgraph(modelpickobj = modelpickobject2)
-rawhistcuts(modelpickobj = modelpickobject2)
+rawuncertgraph(modelpickobj = modelpickobject2, title="", xlab="Optical Density")
+rawdistgraph(modelpickobj = modelpickobject2, title="", xlab="Optical Density")
+rawhistcuts(modelpickobj = modelpickobject2, title="", xlab="Optical Density")
 
-cutoffobject2<-cutoff(modelpickobj = modelpickobject2, cutcomp = 1)
+cutoffobject2<-cutoff(modelpickobj = modelpickobject2, cutcomp = 1, newcertlevel = 0.85)
 
-cutuncertgraph(cutobj = cutoffobject2)
-cutdistgraph(cutobj = cutoffobject2)
+cutuncertgraph(cutobj = cutoffobject2, title="", xlab="Optical Density")
+cutdistgraph(cutobj = cutoffobject2, title="", xlab="Optical Density")
 

### 2017-03-22 Cutoff_Algorithm.R

@@ -64,8 +64,14 @@
 #[IA] Install Packages
 #-----------------------------------------------
 
-install.packages("mixsmsn")
-install.packages("sn")
+#This checks to see if packages are installed and if they are not, then it loads them
+
+if (require("mixsmsn")==F){
+  install.packages("mixsmsn")
+}
+if (require("sn")==F){
+  install.packages("sn")
+}
 
 #-----------------------------------------------
 #[IB] Load Packages
@@ -543,7 +549,7 @@ modelpick<-function(fitobj, dist="", ncomp=NA){
 #[IC4]  rawuncertgraph
 #-----------------------------------------------
 #Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
-rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,title="default",setcolor=c(
+rawuncertgraph<-function(modelpickobj,xlab="x-value",xlim=c(NA,NA),suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
   "turquoise3",      #2
   "royalblue2",      #3
@@ -596,7 +602,7 @@ rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegen
 #[IC5]  rawdistgraph
 #-----------------------------------------------
 #Displays the distributions of all the components for the chosen combination of distribution and number of components. 
-rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,title="default",setcolor=c(
+rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="x-value",setbreaks=100,suppresslegend=F,title="default",setcolor=c(
   "green4",          #1
   "turquoise3",      #2
   "royalblue2",      #3
@@ -694,7 +700,7 @@ rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbrea
 #-----------------------------------------------
 #Displays which component each bar of the histogram would be assigned to. 
 
[…]

### 2017-03-23 Cutoff_Algorithm_170202.R
Delete Cutoff_Algorithm_170202.R

@@ -1,1471 +1,0 @@
-#Master Cutoff Algorithm Functions 
-
-#The cutoff algorithm contained in this R script allows the user to descriminate between a positive and negative sample on a continuous diagnostic without a good gold standard via mixture modeling.
-#The algorithm runs 10 different potential types of mixture models which are based on the normal and skew-normal distributions, looking at 1-5 components each. The optimal distribution and component combination is chosen by BIC (which should be minimized)
-#Note: the fitloops function takes some time (typically 2-3 minutes) to run and can fail, if it does, read the error messages and decide whether to try and run it again, or restrict the maximum number of components tested.
-#The user then checks which distribution has been chosen and can manually chose a different distribution if desired (there should be good scientific or clinical reasoning to choose a different combination)
-#If a model with more than two components is optimal by BIC, then the user must allot components to the positive and negative side using their knowlege of the epidemiological conditions where the data comes from.
-#After the decision of where to cut is made, the user can then output graphs and tables that detail what proportion of the population falls into what category, with various degrees of certainty.
-
-
-#To complete this process, there are 5 necessary steps
-#I)   Install/load packages and load in functions (run lines 73 - 1350)
-#II)  Load in data (run lines: 1370)
-#III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
-#     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
-#     determine which cutpoint is best (if a model with more than two components is optimal) and 3) display the cutoff results
- 
- 
-#-----------------------------------------------
-#Table of Contents
-#-----------------------------------------------
-#[I] Install/load packages and load in functions
-#[IA] Install Packages
-#[IB] Load Packages
-#[IC] Load Functions
[…]
